---
title: 'Simulation Study: General Setting'
output: html_document
date: "2024-10-18"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Libraries
library(data.table)
library(survival)
#install.packages("rtmle")
#devtools::install_github("tagteam/rtmle")
library(rtmle)
library(ggplot2)
library(microbenchmark)
theme_set(theme_bw())
```


# The Simulation Function

We start out by specifying relevant quantities:

```{r}
N <- 100
eta <- c(0.1, 0.1, 0.1, 0.1)
nu <- c(1.1, 1.1, 1.1, 1.1)
# Effect on Operation
beta0 <- c(3, 0, 1, 3)
# Effect on Death
beta1 <- c(0, 0, -1, 1)
# Effect on Censoring
beta2 <- c(3, 0, 1, 1)
# Effect on time varying covariate
beta3 <- c(3, 0, -1, 0.5)
beta <- cbind(beta0, beta1, beta2, beta3)
```

## Simulationg operation data

We can use the function `sim_op_data` to simulate data in the operation setting. 

```{r}
data <- sim_op_data(N = N, beta = beta, eta = eta, nu = nu)
plot_event_data(data) 
```

This corresponds to using the general function `sim_event_data` in the following way.

```{r}
# Checking whether the simulations work
data <- sim_event_data(N = N, beta = beta, eta = eta, nu = nu)
plot_event_data(data) 
```

Model fit

```{r}
# Generating data
N <- 10^4
beta <- matrix(rnorm(16,0,0.75), ncol = 4, nrow = 4)
data_test <- sim_op_data(N = N, beta = beta)

# Transform data into tstart tstop format
data_int <- trans_int_data(data_test)

data_int[, at_risk_oper := as.numeric(A == 0)]
data_int[, at_risk_cov := as.numeric(L == 0)]


# Fit models
survfit_oper <- coxph(Surv(tstart, tstop, Delta == 0) ~ L0 + A0 + L, data = data_int[at_risk_oper == 1], cluster = ID)
survfit_death <- coxph(Surv(tstart, tstop, Delta == 1) ~ L0 + A + A0 + L, data = data_int, cluster = ID)
survfit_cens <- coxph(Surv(tstart, tstop, Delta == 2) ~ L0 + A + A0 + L, data = data_int, cluster = ID)
survfit_cov <- coxph(Surv(tstart, tstop, Delta == 3) ~ L0 + A + A0, data = data_int[at_risk_cov == 1], cluster = ID)
confint(survfit_oper); c(beta[1,1], beta[3,1], beta[4,1])
confint(survfit_death); beta[,2]
confint(survfit_cens); beta[,3]
confint(survfit_cov); beta[1:3,4]

# Visualizing the confidence intervals

CIs <- cbind("Par" = c("oper_L0", "oper_A", "oper_A0", "death_L0", "death_A", "death_A0", "death_L",
                       "cens_L0", "cens_A", "cens_A0", "cens_L","cov_L0", "cov_A", "cov_A0" ),
           rbind(confint(survfit_oper), confint(survfit_death), confint(survfit_cens),confint(survfit_cov)))

rownames(CIs) <- NULL
colnames(CIs) <- c("Par", "Lower", "Upper")
CIs <- data.table(CIs)

CIs[, True_val := c(beta[1,1], beta[3,1], beta[4,1],
               beta[,2],
               beta[,3],
               beta[1:3,4])]

CIs$Lower <- as.numeric(CIs$Lower)
CIs$Upper <- as.numeric(CIs$Upper)


ggplot(data = CIs)+
  geom_point(aes(y = Par, x = Lower))+
  geom_point(aes(y = Par, x = Upper))+
  geom_point(aes(y = Par, x = True_val, col = "Estimate"))+
  scale_color_manual(values = c("Estimate" = "red"))+
  geom_segment(aes(y = Par, yend = Par, x = Lower, xend = Upper))

```

## Simulationg survival data
The function `sim_surv_data` simulates survival data. It is a wrapper of the function `sim_event_data`.

```{r}
data <- sim_surv_data(100)
plot_event_data(data, title = "Survival Data")
```

Model fit

```{r}
beta <- matrix(rnorm(4,0,1), ncol = 2, nrow = 2)
data <- sim_surv_data(N = 6000, beta = beta)

survfit_death <- coxph(Surv(Time, Delta == 1) ~ L0 + A0, data = data)
coefficients(survfit_death); beta[,1]
confint(survfit_death)

survfit_cens <- coxph(Surv(Time, Delta == 2) ~ L0 + A0, data = data)
coefficients(survfit_cens); beta[,2]
confint(survfit_cens)
```

Seems to estimate correctly. 

## Simulating competing risk data

```{r}
data <- sim_cr_data(50)
plot_event_data(data, title = "Competing Risk Data")
```

Model fit

```{r}
beta <- matrix(rnorm(6,0,1), ncol = 3, nrow = 2)
data <- sim_cr_data(N = 10000, beta = beta)

survfit_proc1 <- coxph(Surv(Time, Delta == 0) ~ L0 + A0, data = data)
coefficients(survfit_proc1); beta[,1]
confint(survfit_proc1)

survfit_proc2 <- coxph(Surv(Time, Delta == 1) ~ L0 + A0, data = data)
coefficients(survfit_proc2); beta[,2]
confint(survfit_proc2)

survfit_cens <- coxph(Surv(Time, Delta == 2) ~ L0 + A0, data = data)
coefficients(survfit_cens); beta[,3]
confint(survfit_cens)
```

Also seems to estimate correctly. 

## Profiling

```{r}
library(profvis)

profvis({
  sim_event_data <- function(N,                  # Number of individuals
                             beta = NULL,            # Effects
                             eta = rep(0.1,4),       # Shape parameters
                             nu = rep(1.1,4),        # Scale parameters
                             at_risk = NULL,         # Function defining the setting
                             term_deltas = c(1,2)    # Terminal events
                             )
    {
    ID <- NULL
    if(is.null(beta)){
      beta <- matrix(0, nrow = 4, ncol = 4)
    }
  
  
    if(is.null(at_risk)){
      at_risk <- function(i, L, A) {
        return(c(
          # You are only at risk for an operation if you have not had an operation yet
          as.numeric(A[i] == 0),
          # If you have not died yet or been censored yet, you are at risk for dying or being censored
          1,1,
          # You are only at risk for a change in the covariate process if you have not had a change yet
          as.numeric(L[i] == 0)))
    }}
  
    # Events
    x <- 1:ncol(beta)
  
    # Intensities
    phi <- function(i) {
      exp(L0[i] * beta[1,] + A[i] * beta[2,] + A0[i] * beta[3,] + L[i] * beta[4,])
    }
  
    lambda <- function(t, i) {
      at_risk(i, L, A) * eta * nu * t ^ (nu - 1) * phi(i)
    }
  
    # Summed cumulative hazard
    sum_cum_haz <- function(u, t, i) {
      sum(at_risk(i, L, A) * eta * phi(i) * ((t + u) ^ nu - t ^ nu))
    }
  
    # Inverse summed cumulative hazard function
    inverse_sc_haz <- function(p, t, i, lower_bound = 10^-15, upper_bound = 100) {
      root_function <- function(u) sum_cum_haz(u, t, i) - p
      stats::uniroot(root_function, lower = lower_bound, upper = upper_bound)$root
    }
  
    # Event probabilities
    probs <- function(t, i){
      probs <- lambda(t, i)
      summ <- sum(probs)
      probs / summ
    }
  
    # Draw
    L0 <- stats::runif(N)
    A0 <- stats::rbinom(N, 1, 0.5)
  
    # Initialize
    T_k <- rep(0,N)
    Delta <- -1
    L <- rep(0, N)
    A <- rep(0, N)
    alive <- 1:N
  
    res <- data.table()
  
    while(length(alive) != 0){
      # Simulate time
      V <- stats::runif(N)
      W <- sapply(alive, function(i) inverse_sc_haz(-log(V)[i], T_k[i], i))
      T_k[alive] <- T_k[alive] + W
  
      # Simulate event
      Deltas <- sapply(alive, function(i) sample(x, size = 1, prob = probs(T_k[i], i)) - 1)
  
      kth_event <- data.table(ID = alive,
                 Time = T_k[alive],
                 Delta = Deltas,
                 L0 = L0[alive],
                 L = L[alive],
                 A0 = A0[alive],
                 A = A[alive])
  
      res <- rbind(res, kth_event)
  
      # Update treatment process and covariate change indicators
      A[alive][Deltas == 0] <- 1
      L[alive][Deltas == 3] <- 1
  
      # Who is still alive and uncensored?
      alive <- alive[! Deltas %in% term_deltas]
    }
    setkey(res, ID)
    return(res)
  }
  
  sim_event_data(1000)
})

```


The time is spent on
- Finding the inverse. But I dont think we can find an analytical expression? 
- Using sapply
- Computing the phi function
- Look into implementing parts in C++

# Two Settings

## Setting 1 ("Operation")

```{r}
beta1 <- matrix(ncol = 4, nrow = 4)

# A0 is 0
beta1[3,] <- 0
# L0 øger ssh for L=1 og A=1
beta1[1,c(1,4)] <- c(1,1)
# L=1 øger ssh for A=1
beta1[4,1] <- 1 
# A=1 nedsætter risiko for død
beta1[2,2] <- -1
# L0,L=1 øger risiko for død
beta1[c(1,4),2] <- 1
# Censorering afhænger ikke af noget
beta1[,3] <- 0

# A = 1 spiller ikke nogen rolle for sig selv, siden at begivenheden kun finder sted en enkelt gang
beta1[2,1] <- 0
# På samme vis spiller det ikke nogen rolle hvordan L=1 påvirker risiko for L
beta1[4,4] <- 0
# Min anategelse: A = 1 nedsætter risiko for L=1
beta1[2,4] <- 0.5
```

```{r}
set.seed(2964)
set1_data <- sim_op_data(N = 1000, beta = beta1)
```

```{r}
# Transform data into tstart tstop format
set1_data_int <- trans_int_data(set1_data)

set1_data_int[, at_risk_oper := as.numeric(A == 0)]
set1_data_int[, at_risk_cov := as.numeric(L == 0)]

# Fit models
survfit_oper <- coxph(Surv(tstart, tstop, Delta == 0) ~ L0 + A0 + L, 
                      data = set1_data_int[at_risk_oper == 1], cluster = ID)
survfit_death <- coxph(Surv(tstart, tstop, Delta == 1) ~ L0 + A + A0 + L, 
                       data = set1_data_int, cluster = ID)
survfit_cens <- coxph(Surv(tstart, tstop, Delta == 2) ~ L0 + A + A0 + L, 
                      data = set1_data_int, cluster = ID)
survfit_cov <- coxph(Surv(tstart, tstop, Delta == 3) ~ L0 + A + A0, 
                     data = set1_data_int[at_risk_cov == 1], cluster = ID)

survfit_oper$coefficients; beta1[,1]
survfit_death$coefficients; beta1[,2]
survfit_cens$coefficients; beta1[,3]
survfit_cov$coefficients; beta1[,4]
```


## Setting 2 ("T2D diabetes")

```{r}
beta2 <- matrix(ncol = 4, nrow = 4)

# Ingen A
beta2[2,] <- 0
beta2[,1] <- 0
# L0 øger ssh for L=1
beta2[1,4] <- 1
# A0=1 mindsker ssh for L=1
beta2[3,4] <- -1
# L0 øger risiko for død
beta2[1,2] <- 1
# A0 nedsætter risiko for død
beta2[3,2] <- -1
# Det ikke nogen rolle hvordan L=1 påvirker risiko for L
beta2[4,4] <- 0

# Censorering afhænger ikke af noget (NB. min antagelse)
beta2[,3] <- 0

# 2a: L=1 sætter risiko for død op 
beta2a <- beta2
beta2a[4,2] <- 1

# 2b: risiko for død afhænger ikke af L
beta2b <- beta2
beta2b[4,2] <- 0
```

```{r}
set.seed(2964)
set2a_data <- sim_op_data(N = 1000, beta = beta2a)
set2b_data <- sim_op_data(N = 1000, beta = beta2b)
```


```{r}
# Transform data into tstart tstop format
set2a_data_int <- trans_int_data(set2a_data)

set2a_data_int[, at_risk_cov := as.numeric(L == 0)]

# Fit models
survfit_death <- coxph(Surv(tstart, tstop, Delta == 1) ~ L0  + A0 + L, 
                       data = set2a_data_int, cluster = ID)
survfit_cens <- coxph(Surv(tstart, tstop, Delta == 2) ~ L0 + A0 + L, 
                      data = set2a_data_int, cluster = ID)
survfit_cov <- coxph(Surv(tstart, tstop, Delta == 3) ~ L0 + A0, 
                     data = set2a_data_int[at_risk_cov == 1], cluster = ID)

survfit_death$coefficients; beta2a[,2]
survfit_cens$coefficients; beta2a[,3]
survfit_cov$coefficients; beta2a[,4]
```

```{r}
# Transform data into tstart tstop format
set2b_data_int <- trans_int_data(set2b_data)

set2b_data_int[, at_risk_cov := as.numeric(L == 0)]

# Fit models
survfit_death <- coxph(Surv(tstart, tstop, Delta == 1) ~ L0  + A0 + L, 
                       data = set2b_data_int, cluster = ID)
survfit_cens <- coxph(Surv(tstart, tstop, Delta == 2) ~ L0 + A0 + L, 
                      data = set2b_data_int, cluster = ID)
survfit_cov <- coxph(Surv(tstart, tstop, Delta == 3) ~ L0 + A0, 
                     data = set2b_data_int[at_risk_cov == 1], cluster = ID)

survfit_death$coefficients; beta2b[,2]
survfit_cens$coefficients; beta2b[,3]
survfit_cov$coefficients; beta2b[,4]
```


Det hele ser super ud!

# Interventions

1. Simulate

```{r}
N <- 10000
beta <- matrix(rnorm(16,0,0.75), ncol = 4, nrow = 4)
data_int <- sim_op_data(N = N, beta = beta)
```

2. Fit a Weibull 

```{r}
data_int[, k:= ave(ID, ID, FUN = seq_along) - 1]
survfit_op <- survreg(Surv(Time, Delta == 0) ~ L0 + A0 + L, data = data_int, dist='weibull')
survfit_death <- survreg(Surv(Time, Delta == 1) ~ L0 + A + A0 + L, data = data_int, dist='weibull')
survfit_cens <- survreg(Surv(Time, Delta == 2) ~ L0 + A + A0 + L, data = data_int, dist='weibull')
survfit_cov <- survreg(Surv(Time, Delta == 3) ~ L0 + A + A0, data = data_int, dist='weibull')
```

3. Estimate parameters

```{r}
shape_est <- c(1/survfit_op$scale, 1/survfit_death$scale, 1/survfit_cens$scale, 1/survfit_cov$scale) 
shape_est; 1.1
scale_est <- c(1/exp(survfit_op$coefficients[1]), 1/exp(survfit_death$coefficients[1]), 
               1/exp(survfit_cens$coefficients[1]), 1/exp(survfit_cov$coefficients[1])) 
scale_est; 0.1
```
4. Simulate new data using the estimated parameters - without censoring

The without censoring, can be implemented in two ways, setting nu/eta = 0 for the censorign process or modifying the at risk function for censoring.

A possible intervention would be to increase the number of operations. Here we do this by adding the constant 1 to all beta coefficients in the operation intensity. 

```{r}
at_risk_int <- function(i, L, A) {return(c(
  # You are only at risk for an operation if you have not had an operation yet
  as.numeric(A[i] == 0),
  # If you have not died yet you are at risk for dying
  1,
  # In the intervention scenario, you cannot be censored
  0,
  # You are only at risk for a change in the covariate process if you have not had a change yet
  as.numeric(L[i] == 0)))}

beta_int <- beta
# Increasing the number of operations
beta_int[,1] <- beta_int[,1] + 1

sim_int <- sim_event_data(N = 1000, beta = beta_int, eta = scale_est, nu = shape_est, at_risk = at_risk_int)
plot_event_data(sim_int)
```

5. Estimate causal effects

To do: implement correctly!

```{r}
mean(data_int[data_int$Delta == 0,]$Time) - mean(sim_int[sim_int$Delta == 0,]$Time)
# You are under operation sooner
mean(data_int[data_int$Delta == 1,]$Time) - mean(sim_int[sim_int$Delta == 1,]$Time)
# You die sooner
mean(data_int[data_int$Delta == 3,]$Time) - mean(sim_int[sim_int$Delta == 3,]$Time)
# There is a change in your covariate process sooner
```

. 
