---
title: 'Simulation Study: Two Settings'
output: html_document
date: "2024-10-18"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Libraries
library(data.table)
library(survival)
#install.packages("rtmle")
devtools::install_github("tagteam/rtmle")
library(rtmle)
library(ggplot2)
library(microbenchmark)
theme_set(theme_bw())
```

# Two Settings

## Setting 1 ("Operation")

```{r}
set.seed(2964)
set1_data <- sim_data_setting1(N = 5000)
set1_data_fu <- sim_data_setting1(N = 5000, followup = 5)
plot_event_data(set1_data)
plot_event_data(set1_data_fu)
```

## Setting 2 ("T2D diabetes")

```{r}
set.seed(2964)
set2a_data <- sim_data_setting2(N = 10000, setting = "a")
set2b_data <- sim_data_setting2(N = 10000, setting = "b")
set2a_fu_data <- sim_data_setting2(N = 10000, setting = "a", followup = 5)
set2b_fu_data <- sim_data_setting2(N = 10000, setting = "b", followup = 5)
```


```{r}
plot_event_data(set2a_data)
plot_event_data(set2a_fu_data)
```


### Proportions

```{r}
summary_sim_event <- function(data) {
  N <- nrow(data)
  prop_op <- nrow(data[Delta == 0])/N
  prop_death <- nrow(data[Delta == 1])/N
  prop_cens <- nrow(data[Delta == 2])/N
  prop_fu_cens <- nrow(data[Delta == "C"])/N
  prop_cov <-  nrow(data[Delta == 3])/N
  
  mean_T <- mean(data$Time)
  max_T <- max(data$Time)
  
  return(c("Prop Op" = prop_op, "Prop Death" = prop_death, "Prop Cens" = prop_cens, 
                    "Prop FU Cens" = prop_fu_cens, "Prop Cov" = prop_cov, "Mean Time" = mean_T, 
                    "Max Time" = max_T))
}
```

```{r}
summary_sim_event(set1_data)
summary_sim_event(set1_data_fu)
summary_sim_event(set2a_data)
summary_sim_event(set2a_fu_data)
summary_sim_event(set2b_data)
summary_sim_event(set2b_fu_data)
```

### Fitting models


```{r}
# Transform data into tstart tstop format
set2a_data_int <- trans_int_data(set2a_data)

set2a_data_int[, at_risk_cov := as.numeric(L == 0)]

# Fit models
survfit_death <- coxph(Surv(tstart, tstop, Delta == 1) ~ L0  + A0 + L, 
                       data = set2a_data_int, cluster = ID)
survfit_cens <- coxph(Surv(tstart, tstop, Delta == 2) ~ L0 + A0 + L, 
                      data = set2a_data_int, cluster = ID)
survfit_cov <- coxph(Surv(tstart, tstop, Delta == 3) ~ L0 + A0, 
                     data = set2a_data_int[at_risk_cov == 1], cluster = ID)

survfit_death$coefficients 
survfit_cens$coefficients 
survfit_cov$coefficients
```

```{r}
# Transform data into tstart tstop format
set2b_data_int <- trans_int_data(set2b_data)

set2b_data_int[, at_risk_cov := as.numeric(L == 0)]

# Fit models
survfit_death <- coxph(Surv(tstart, tstop, Delta == 1) ~ L0  + A0 + L, 
                       data = set2b_data_int, cluster = ID)
survfit_cens <- coxph(Surv(tstart, tstop, Delta == 2) ~ L0 + A0 + L, 
                      data = set2b_data_int, cluster = ID)
survfit_cov <- coxph(Surv(tstart, tstop, Delta == 3) ~ L0 + A0, 
                     data = set2b_data_int[at_risk_cov == 1], cluster = ID)

survfit_death$coefficients 
survfit_cens$coefficients 
survfit_cov$coefficients
```


### Proportions

Note that we can affect the intensities by:

$$\lambda^x_k(t\, \vert \, \mathcal{F}_{t-}) =  R^x(t,k)  \eta^x \nu^x t^{\nu^x - 1} 
 \exp(\beta^T Z)$$

changing the value of $\eta^x$. For a follow up time $K$ 

```{r}
eta_vals <- seq(0.05, 0.35, by = 0.1)
eta_vec <- expand.grid(eta1 = eta_vals, eta2 = 0.1, eta3 = 0.1, eta4 = eta_vals)

K <- 5
summaries <- matrix(ncol = 7, nrow = nrow(eta_vec))

for(i in 1:nrow(eta_vec)){
  print(i)
  eta <- eta_vec[i,]
  sim <- sim_data_setting1(100, eta = eta, followup = K)
  summaries[i,] <- summary_sim_event(sim)
}

colnames(summaries) <- names(summary_sim_event(sim))
summaries
```

### Investigating model misspecifications

Function for fitting and extracting model estimates, when the model is specified correctly, misspecified in terms of missing time varying covariate effect and misspecified in terms of missing operation effect.  

```{r}
fit_mod_set1 <- function(data){
  trans_data <- trans_int_data(data)
  
  trans_data[, at_risk_oper := as.numeric(A == 0)]
  trans_data[, at_risk_cov := as.numeric(L == 0)]
  
  results <- matrix(nrow = 10, ncol = 3)
  colnames(results) <- c("True", "-L", "-A")

  # True models:
  survfit_oper <- coxph(Surv(tstart, tstop, Delta == 0) ~ L0 + L, 
                        data = trans_data[at_risk_oper == 1], cluster = ID)
  survfit_death <- coxph(Surv(tstart, tstop, Delta == 1) ~ L0 + A + L, 
                        data = trans_data, cluster = ID)
  survfit_cens <- coxph(Surv(tstart, tstop, Delta == 2) ~ L0 + A + L, 
                        data = trans_data, cluster = ID)
  survfit_cov <- coxph(Surv(tstart, tstop, Delta == 3) ~ L0 + A, 
                      data = trans_data[at_risk_cov == 1], cluster = ID)
  
  results[,1] <- c(survfit_oper$coefficients, survfit_death$coefficients, 
                   survfit_cens$coefficients, survfit_cov$coefficients)
  
  # Cox models only with time varying covariate:

  survfit_oper <- coxph(Surv(tstart, tstop, Delta == 0) ~ L0 + L, 
                        data = set1_data_int[at_risk_oper == 1], cluster = ID)
  survfit_death <- coxph(Surv(tstart, tstop, Delta == 1) ~ L0 + L, 
                         data = set1_data_int, cluster = ID)
  survfit_cens <- coxph(Surv(tstart, tstop, Delta == 2) ~ L0 + L, 
                        data = set1_data_int, cluster = ID)
  survfit_cov <- coxph(Surv(tstart, tstop, Delta == 3) ~ L0, 
                       data = set1_data_int[at_risk_cov == 1], cluster = ID)
  
  results[,2] <- c(survfit_oper$coefficients, survfit_death$coefficients[1], NA, 
                   survfit_death$coefficients[2], survfit_cens$coefficients[1], NA, 
                   survfit_cens$coefficients[2], survfit_cov$coefficients, NA)

  # Cox models only with operation:
  survfit_oper <- coxph(Surv(tstart, tstop, Delta == 0) ~ L0, 
                        data = set1_data_int[at_risk_oper == 1], cluster = ID)
  survfit_death <- coxph(Surv(tstart, tstop, Delta == 1) ~ L0 + A, 
                         data = set1_data_int, cluster = ID)
  survfit_cens <- coxph(Surv(tstart, tstop, Delta == 2) ~ L0 + A, 
                        data = set1_data_int, cluster = ID)
  survfit_cov <- coxph(Surv(tstart, tstop, Delta == 3) ~ L0 + A, 
                       data = set1_data_int[at_risk_cov == 1], cluster = ID)
  
  results[,3] <- c(survfit_oper$coefficients, NA, survfit_death$coefficients, NA,
                   survfit_cens$coefficients, NA, survfit_cov$coefficients)
  
  return(results)
  
}

fit_mod_set1(set1_data)
```

Initial investigations:

```{r}
# Transform data into tstart tstop format
set1_data_int <- trans_int_data(set1_data)

set1_data_int[, at_risk_oper := as.numeric(A == 0)]
set1_data_int[, at_risk_cov := as.numeric(L == 0)]

# True models:
survfit_oper <- coxph(Surv(tstart, tstop, Delta == 0) ~ L0 + L, 
                      data = set1_data_int[at_risk_oper == 1], cluster = ID)
survfit_death <- coxph(Surv(tstart, tstop, Delta == 1) ~ L0 + A + L, 
                       data = set1_data_int, cluster = ID)
survfit_cens <- coxph(Surv(tstart, tstop, Delta == 2) ~ L0 + A + L, 
                      data = set1_data_int, cluster = ID)
survfit_cov <- coxph(Surv(tstart, tstop, Delta == 3) ~ L0 + A, 
                     data = set1_data_int[at_risk_cov == 1], cluster = ID)

# Cox models only with time varying covariate:

survfit_oper1 <- coxph(Surv(tstart, tstop, Delta == 0) ~ L0 + L, 
                      data = set1_data_int[at_risk_oper == 1], cluster = ID)
survfit_death1 <- coxph(Surv(tstart, tstop, Delta == 1) ~ L0 + L, 
                       data = set1_data_int, cluster = ID)
survfit_cens1 <- coxph(Surv(tstart, tstop, Delta == 2) ~ L0 + L, 
                      data = set1_data_int, cluster = ID)
survfit_cov1 <- coxph(Surv(tstart, tstop, Delta == 3) ~ L0, 
                     data = set1_data_int[at_risk_cov == 1], cluster = ID)

# Cox models only with operation:
survfit_oper2 <- coxph(Surv(tstart, tstop, Delta == 0) ~ L0, 
                      data = set1_data_int[at_risk_oper == 1], cluster = ID)
survfit_death2 <- coxph(Surv(tstart, tstop, Delta == 1) ~ L0 + A, 
                       data = set1_data_int, cluster = ID)
survfit_cens2 <- coxph(Surv(tstart, tstop, Delta == 2) ~ L0 + A, 
                      data = set1_data_int, cluster = ID)
survfit_cov2 <- coxph(Surv(tstart, tstop, Delta == 3) ~ L0 + A, 
                     data = set1_data_int[at_risk_cov == 1], cluster = ID)


survfit_oper$coefficients; survfit_oper1$coefficients; survfit_oper2$coefficients
survfit_death$coefficients; survfit_death1$coefficients; survfit_death2$coefficients
survfit_cens$coefficients; survfit_cens1$coefficients; survfit_cens2$coefficients
survfit_cov$coefficients; survfit_cov1$coefficients; survfit_cov2$coefficients

```

More systematic investigation:

```{r}
set.seed(9458)
B <- 100
N <- 500

for(i in 1:B){
  sim <- sim_data_setting1(N = 1000)
  
  
  
}
```


































