---
title: "Estimations"
author: "Michaela Lukacova"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, include=FALSE}
# Libraries
library(data.table)
library(survival)
library(simevent)
library(ggplot2)
library(prodlim)
library(rstpm2)
theme_set(theme_bw())
```

# T2D

## Generate data

```{r}
set.seed(373)
data <- simT2D(N = 5000, beta_A0_L = -1)
```

## Estimation

We fit the models

```{r}
# Transform data into tstart tstop format
data_int <- IntFormatData(data, N_cols = 6)

# Fit Cox models
cox_death <- coxph(Surv(tstart, tstop, Delta == 1) ~ L0 + A0 + L, data = data_int)
cox_t2d <- coxph(Surv(tstart, tstop, Delta == 2) ~ L0 + A0, data = data_int[L == 0])
```

A little test

```{r}
predict_data <- data.frame("A0" = c(1,1), "L0" = c(0.5, 0.5), "L" = c(0,1))

basehazz_t2d <- basehaz(cox_t2d, centered = TRUE)
basehazz_death <- basehaz(cox_death, centered = TRUE)
cox_term_t2d <- exp(predict(cox_t2d, newdata=predict_data, type="lp"))
cox_term_death <- exp(predict(cox_death, newdata=predict_data, type="lp"))
cum_int_t2d <- outer(basehazz_t2d[['hazard']], cox_term_t2d, "*")
cum_int_death <- outer(basehazz_death[['hazard']], cox_term_death, "*")

true_surv_death_func1 <- function(t) exp( - 0.1  * t^(1.1) * exp(1 * 0.5 + 0 * (-1) + 0 * 1))
true_surv_death_func2 <- function(t) exp( - 0.1  * t^(1.1) * exp(1 * 0.5 + 0 * (-1) + 1 * 1))
true_surv_t2d_func <- function(t) exp( - 0.1  * t^(1.1) * exp(1 * 0.5 + -1 * 1))

gridExtra::grid.arrange(
  ggplot(data = data.frame(x = basehazz_death$time, y = exp(-cum_int_death[,1])))+
    geom_line(aes(x = x, y = y), color = "red") +
    geom_function(fun = true_surv_death_func1),
  ggplot(data = data.frame(x = basehazz_death$time, y = exp(-cum_int_death[,2])))+
    geom_line(aes(x = x, y = y), color = "red") +
    geom_function(fun = true_surv_death_func2),
  ggplot(data = data.frame(x = basehazz_t2d$time, y = exp(-cum_int_t2d[,1])))+
    geom_line(aes(x = x, y = y), color = "red") +
    geom_function(fun = true_surv_t2d_func),
  nrow = 2)
```

Estimates quite well. 

## Simulation function

In order to simulate new data from an observed data set, we start out by initializing $N$ observations. We draw the $A_0$ covariate from a Bernoulli with probability $0.5$  and the $L_0$ we sample from the empirical distribution. We initialize the $T_0$ as $0$ and the cumulative intensity at this time as $0$. 

Then as long as we have individuals that are alive, we for each individual find the cumulative intensity functions of death and T2D. In order to simulate the next event time $T_k = T_{k-1} + W_k$ we need to find the waiting time $W_k$. We find this time by finding the waiting time of each event and taking the minimum, in this case, the waiting time of the death event $W_k^{death}$ and the waiting time of the T2D event $W_k^{T2D}$. $W_k^x$ is distributed with cumulative intensity
$$ \tilde{\Lambda}^x(u|T_{k-1} = t) = \Lambda^x(u + t ) - \Lambda^x(t)$$
We then have that
$$(\tilde{\Lambda}^x)^{-1}(-\log(U)) \sim \tilde{\Lambda}^x(u)$$
Where $U$ is a uniform random variable. The inverse cumulative intensity of the waiting times is given by
$$(\tilde{\Lambda}^x)^{-1}(u |T_{k-1} = t, \mathcal{F}_{(u+t)-}) = \Lambda^{-1}_x(-\log u + \Lambda(t|F_{t-})|\mathcal{F}_{(u+t)-}) - t$$
So in order to simulate event specific waiting times, we draw uniform random variables take the minus log and of this number take the inverse cumulative intensity and subtract $t$. 

The simulation procedure is implemented in the function `simEventDataCox` below. 

```{r}
simEventDataCox <- function(N, 
                            cox_fits,
                            L0_old, 
                            max_events = 3, 
                            proc_names = c("D", "L"),
                            n_event_max = c(1,1),
                            term_events = 1) {
  
  # Initialize
  num_events <- length(cox_fits)                          # Number of events
  alive <- 1:N                                            # Vector for keeping track of who is alive
  num_alive <- N                                          # Number of alive individuals
  T_k <- rep(0, N)                                        # Last event time
  
  # Matrix for storing data
  sim_data <- matrix(nrow = N, ncol = (2 + num_events))
  colnames(sim_data) <- c("L0", "A0", proc_names)
  sim_data[,1] <- sample(L0_old, N, replace = TRUE)       # sample with replacement from old data
  sim_data[,2] <- rbinom(N, 1, 0.5)                       # A0 
  sim_data[,3:(num_events+2)] <- 0
  sim_data <- data.frame(sim_data)
  
  # List for results
  res_list <- vector("list", max_events)                  # For results
  idx <- 1                                                # Index
  
  # Function to find the cumulative intensity 
  cum_int <- function(int_vals, time_vals, t) {
    cum_int_t <- int_vals[time_vals <= t]                 # All the cum.int. values at times smaller than t
    ifelse(length(cum_int_t) == 0, 0, max(cum_int_t))     # We take the max of these, or if there are none return 0
  }
  
  # Function to find the inverse cumulative intensity
  inv_cum_int <- function(int_vals, time_vals, p) {
    T_k_1 <- time_vals[int_vals >= p]                      # We find the times for which intensity > p
    ifelse(length(T_k_1) == 0, max(time_vals), min(T_k_1))  # We find the inverse cumulative intensity
  }
  
  # Base hazard
  basehazz_list <- lapply(cox_fits, function(model) basehaz(model, centered = FALSE))
  # Jump times of the base hazard
  jump_times <- lapply(basehazz_list, function(df) df[["time"]])
  basehazz_list <- lapply(basehazz_list, function(df) df[["hazard"]])
  
  while(length(alive) != 0){
    # Calculate the Cox term
    cox_term <- lapply(cox_fits, function(model) exp(predict(model, newdata = sim_data, type="lp")))
    
    # Calculate the cumulative intensity
    cumInt_list <- Map(function(base_vec, cox_vec) {
      outer(base_vec, cox_vec, "*")
      }, basehazz_list, cox_term)
    
    # We find the cumulative intensity of event j and the ith individual at the event time T_k
    cum_int_Tk <- matrix(0, ncol = num_events, nrow = num_alive)
    for(j in 1:num_events){
      cum_int_Tk[,j] <- sapply(1:num_alive, function(i) cum_int(cumInt_list[[j]][,i], jump_times[[j]], T_k[i]))
    }
  
    # Simulate the uniform random variable
    U <- -log(runif(num_alive))
    V <- apply(cum_int_Tk, 2, function(cumInt) cumInt + U)
    # Find the event times
    event_times <- matrix(ncol = num_events, nrow = num_alive)
    for(j in 1:num_events){
      event_times[,j] <- sapply(1:num_alive, function(i) inv_cum_int(cumInt_list[[j]][,i], jump_times[[j]], V[i,j]))
    }
    
    # Checking whether the true inverse cumulative intensity corresponds to the one we calculate
    # It does!!!!!!!!!
    inv_cum_int_true <- matrix(ncol = num_events, nrow = num_alive)
    for(j in 1:num_events){
      inv_cum_int_true[,j] <- sapply(1:num_alive, function(i) true_inv_cum_int(V[i,j]))
    }
    # I seriously dont get it...
    
    # How many times can you experience the various events?
    for(j in 1:num_events){
      event_times[sim_data[, (2+j)] == n_event_max[j], j] <- Inf
    }
    
    # The next event is the minimum of these events
    T_k <- apply(event_times, 1, min)
    Deltas <- apply(event_times, 1, which.min)
    
    # Update event counts
    for(i in 1:num_alive){
      sim_data[i, (Deltas[i] + 2)] <- sim_data[i, (Deltas[i] + 2)] + 1
    }
    
    # Store data
    kth_event <- data.table(ID = alive,
                            Time = T_k,
                            Delta = Deltas)
  
    res_list[[idx]] <- cbind(kth_event, data.table::as.data.table(sim_data))
    idx <- idx + 1
    
    # Who is still alive?
    alive <- alive[!(Deltas %in% term_events)]
    num_alive <- length(alive)
    # For the next iteration we only keep data from individuals alive
    T_k <- T_k[alive] 
    sim_data <- sim_data[alive, , drop = FALSE]
  }
  
  res <- data.table::rbindlist(res_list)
  setkey(res, ID)
  return(res)
}
```

## Simulate new data

```{r}
cox_fits <- list(death = cox_death, t2d = cox_t2d)
new_data <- simEventDataCox(3000, cox_fits, L0_old = data$L0)
```


## check whether simulated data has the desired intensitities...

```{r}
# Transform data into tstart tstop format
data_int2 <- IntFormatData(new_data, N_cols = 6:7)

# Fit Cox models
cox_death2 <- coxph(Surv(tstart, tstop, Delta == 1) ~ L0 + A0 + L, data = data_int2)
cox_t2d2 <- coxph(Surv(tstart, tstop, Delta == 2) ~ L0 + A0, data = data_int2[L == 0])
basehazz_t2d2 <- basehaz(cox_t2d2, centered = TRUE)
basehazz_death2 <- basehaz(cox_death2, centered = TRUE)

predict_data <- data.frame("A0" = c(1,1), "L0" = c(0.5, 0.5), "L" = c(0,1))
cox_term_t2d2 <- exp(predict(cox_t2d2, newdata=predict_data, type="lp"))
cox_term_death2 <- exp(predict(cox_death2, newdata=predict_data, type="lp"))
cum_int_t2d2 <- outer(basehazz_t2d2[['hazard']], cox_term_t2d2, "*")
cum_int_death2 <- outer(basehazz_death2[['hazard']], cox_term_death2, "*")

gridExtra::grid.arrange(
  ggplot()+
    geom_line(aes(x = basehazz_death$time, y = exp(-cum_int_death[,1])), color = "red") +
    geom_line(aes(x = basehazz_death2$time, y = exp(-cum_int_death2[,1])), color = "blue") + 
    geom_function(fun = true_surv_death_func1),
  ggplot()+
    geom_line(aes(x = basehazz_death$time, y = exp(-cum_int_death[,2])), color = "red") +
    geom_line(aes(x = basehazz_death2$time, y = exp(-cum_int_death2[,2])), color = "blue") + 
    geom_function(fun = true_surv_death_func2),
  ggplot()+
    geom_line(aes(x = basehazz_t2d$time, y = exp(-cum_int_t2d[,1])), color = "red") +
    geom_line(aes(x = basehazz_t2d2$time, y = exp(-cum_int_t2d2[,1])), color = "blue") + 
    geom_function(fun = true_surv_t2d_func),
  nrow = 2)
```

## Trying with competing risk data

```{r}
set.seed(373)
data <- simCRdata(10000)
predict_data <- data.frame("A0" = 0, "L0" = 0.5)
true_surv_death_func <- function(t) exp( - 0.1  * t^(1.1)) 
true_surv_t2d_func <- function(t) exp( - 0.1  * t^(1.1))

# Fit Cox models
cox_death <- coxph(Surv(Time, Delta == 1) ~ L0 + A0, data = data)
cox_t2d <- coxph(Surv(Time, Delta == 2) ~ L0 + A0, data = data)
surv_death <- survfit(cox_death, newdata = predict_data, method = "breslow")
surv_t2d <- survfit(cox_t2d, newdata = predict_data, method = "breslow")

cox_fits <- list(death = cox_death, t2d = cox_t2d)
new_data <- simEventDataCox(5000, cox_fits, L0_old = data$L0, term_events = c(1,2))

# Fit Cox models
cox_death2 <- coxph(Surv(Time, Delta == 1) ~ L0 + A0, data = new_data)
cox_t2d2 <- coxph(Surv(Time, Delta == 2) ~ L0 + A0, data = new_data)

surv_death2 <- survfit(cox_death2, newdata = predict_data, method = "breslow")
surv_t2d2 <- survfit(cox_t2d2, newdata = predict_data, method = "breslow")

ggplot()+
  geom_line(aes(x = surv_death$time, y = surv_death$surv), color = "red") +
  geom_line(aes(x = surv_death2$time, y = surv_death2$surv), color = "blue") +
  geom_function(fun = true_surv_death_func)

ggplot()+
  geom_line(aes(x = surv_t2d$time, y = surv_t2d$surv), color = "red") +
  geom_line(aes(x = surv_t2d2$time, y = surv_t2d2$surv), color = "blue") +
  geom_function(fun = true_surv_t2d_func)
```

