---
title: "Kaplan-Meier Simulation Study"
output:
  pdf_document: default
  html_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, include=FALSE}
# Libraries
library(data.table)
library(survival)
library(simevent)
library(ggplot2)
library(microbenchmark)
theme_set(theme_bw())
```

# Aim

In the following a simulation study is conducted, where we investigate the Kaplan-Meier estimator. The estimand is
$$S(\tau | A_0 = 1) = P(T \geq \tau | A_0 = 1) , \quad S(\tau | A_0 = 0) = P(T \geq \tau | A_0 = 0)$$
For some $\tau$. $T$ is here the survival time *post* T2D diagnose. As mentioned the survival function will be estimated by the Kaplan-Meier estimator. Post T2D diagnose the setting is simpler, as there are no competing events, only death and censoring. The Kaplan-Meier estimator will then be an unbiased estimator of the survival probability. 

The true value of $P(T \geq \tau | A_0 = a_0)$ can be estimated by a Monte Carlo approach. That is by simulating a large data set without censoring, and calculating the proportion of dead at time $\tau$ post T2D diagnose, in respectively treatment and placebo group. 

After having estimated the true value, we generate $B$ data sets with $N$ individuals. For each data set we find the Kaplan-Meier estimate of $P(T \geq \tau | A_0 = a_0)$, the SE, and the upper and lower confidence bands. We plot a histogram of the estimates, calculate the bias of the estimator and the coverage of the associated CI. 

We conduct the simulation study for the following 4 different parameter values

```{r}
beta_L0_L <- c(1.5, 1.3, 2, 0.9)
beta_A0_L <- c(-1, -2, -3, -1.2)
beta_L_D <- c(0.9, 1, 0.5, 0.8)
beta_L0_D <- c(0.6, 0.7, 0, 0.6)
beta_A0_D <- c(0, 0, -0.1, -0.2)
```

And the following values of $\eta$ and $\nu$

```{r}
eta <- c(0.1,0.3,0.1)
nu <- c(1.1,1.3,1.1)
```

# True estimate

We simulate the large data

```{r}
set.seed(836)
N <- 2*10^5
data_list <- list()

for(i in 0:9){
  print(i)
  data_list[[i*4 + 1]] <- simT2D(N = N, eta = eta, nu = nu, beta_L0_D = beta_L0_D[1], 
                beta_L0_L = beta_L0_L[1], beta_A0_L = beta_A0_L[1], 
                beta_L_D = beta_L_D[1], beta_A0_D = beta_A0_D[1], cens = 0)
  data_list[[i*4 + 2]] <- simT2D(N = N, eta = eta, nu = nu, beta_L0_D = beta_L0_D[2], 
                beta_L0_L = beta_L0_L[2], beta_A0_L = beta_A0_L[2], 
                beta_L_D = beta_L_D[2], beta_A0_D = beta_A0_D[2], cens = 0)
  data_list[[i*4 + 3]] <- simT2D(N = N, eta = eta, nu = nu, beta_L0_D = beta_L0_D[3], 
                beta_L0_L = beta_L0_L[3], beta_A0_L = beta_A0_L[3], 
                beta_L_D = beta_L_D[3], beta_A0_D = beta_A0_D[3], cens = 0)
  data_list[[i*4 + 4]] <- simT2D(N = N, eta = eta, nu = nu,, beta_L0_D = beta_L0_D[4], 
                beta_L0_L = beta_L0_L[4], beta_A0_L = beta_A0_L[4], 
                beta_L_D = beta_L_D[4], beta_A0_D = beta_A0_D[4], cens = 0)
}
```

We calculate the survival proportion $\tau = 2$ years post T2D diagnose for each data set. 

```{r}
tau <- 1
surv_prop1 <- numeric(16)
surv_prop0 <- numeric(16)

for(i in 1:16){
  # T2D events
  T2D_events <- data_list[[i]][Delta == 2]

  # T2D people
  T2D_peeps <- data_list[[i]][ID %in% T2D_events$ID]
  
  # Setting T_0 to debut time of diabetes
  T2D_peeps[, Time_T2D := Time - min(Time), by = ID]
  
  # Removing the new Time 0
  T2D_peeps <- T2D_peeps[Delta != 2]
  
  surv_prop1[i] <- nrow(T2D_peeps[Time_T2D > tau & A0 == 1]) / nrow(T2D_peeps[A0 == 1]) 
  surv_prop0[i] <- nrow(T2D_peeps[Time_T2D > tau & A0 == 0]) / nrow(T2D_peeps[A0 == 0]) 
}
```

We find the mean and sd

```{r}
c <- numeric(4)
surv_prop_mean0 <- numeric(4)
surv_prop_sd1 <- numeric(4)
surv_prop_sd0 <- numeric(4)

for(i in 1:4){
  surv_prop_mean1[i] <- mean(surv_prop1[0 : 3 * 4 + i])
  surv_prop_mean0[i] <- mean(surv_prop0[0 : 3 * 4 + i])
  surv_prop_sd1[i] <- sd(surv_prop1[0 : 3 * 4 + i])
  surv_prop_sd0[i] <- sd(surv_prop0[0 : 3 * 4 + i])
}

```

# Kaplan-Meier simulation study 

For each of set of parameter values, we simulate $B$ data sets with $N$ subjects, and calculate the KM estimate and CI's, we do this with use of the function `simStudyT2D`. 

```{r}
set.seed(836)
N <- 1000
B <- 500
res1 <- simStudyT2D(N = N, B = B, eta = eta, nu = nu, beta_L0_D = beta_L0_D[1], 
                beta_L0_L = beta_L0_L[1], beta_A0_L = beta_A0_L[1], 
                beta_L_D = beta_L_D[1], beta_A0_D = beta_A0_D[1], tau = tau)
res2 <- simStudyT2D(N = N, B = B, eta = eta, nu = nu, beta_L0_D = beta_L0_D[2], 
                beta_L0_L = beta_L0_L[2], beta_A0_L = beta_A0_L[2], 
                beta_L_D = beta_L_D[2], beta_A0_D = beta_A0_D[2], tau = tau)
res3 <- simStudyT2D(N = N, B = B, eta = eta, nu = nu, beta_L0_D = beta_L0_D[3], 
                beta_L0_L = beta_L0_L[3], beta_A0_L = beta_A0_L[3], 
                beta_L_D = beta_L_D[3], beta_A0_D = beta_A0_D[3], tau = tau)
res4 <- simStudyT2D(N = N, B = B, eta = eta, nu = nu, beta_L0_D = beta_L0_D[4], 
                beta_L0_L = beta_L0_L[4], beta_A0_L = beta_A0_L[4], 
                beta_L_D = beta_L_D[4], beta_A0_D = beta_A0_D[4], tau = tau)
```

## Histograms of estimate

```{r}
plot_data <- data.table(ests = c(res1[,'Est 0'], res1[,'Est 1'], 
                                 res2[,'Est 0'], res2[,'Est 1'],
                                 res3[,'Est 0'], res3[,'Est 1'],
                                 res4[,'Est 0'], res4[,'Est 1']),
                        Lower = c(res1[,'Lower 0'], res1[,'Lower 1'], 
                                 res2[,'Lower 0'], res2[,'Lower 1'],
                                 res3[,'Lower 0'], res3[,'Lower 1'],
                                 res4[,'Lower 0'], res4[,'Lower 1']),
                        Upper = c(res1[,'Upper 0'], res1[,'Upper 1'], 
                                 res2[,'Upper 0'], res2[,'Upper 1'],
                                 res3[,'Upper 0'], res3[,'Upper 1'],
                                 res4[,'Upper 0'], res4[,'Upper 1']),
                        A0 = rep(c(rep(0, B), rep(1, B)),4),
                        Setting = c(rep("A", 2 * B), rep("B", 2 * B), 
                                    rep("C", 2 * B),rep("D", 2 * B)),
                        true_val = c(rep(surv_prop0[1], B), rep(surv_prop1[1],B),
                                     rep(surv_prop0[2], B), rep(surv_prop1[2],B),
                                     rep(surv_prop0[3], B), rep(surv_prop1[3],B),
                                     rep(surv_prop0[4], B), rep(surv_prop1[4],B)))
pp <- ggplot(plot_data) +
  geom_histogram(aes(x = ests, y = ..density..), color = "white", fill = "steelblue")+
  geom_vline(aes(xintercept = true_val), color = "darkred")+
  facet_wrap( ~ A0 + Setting, ncol = 4)

ggsave("hist_sim_KM.jpeg", pp, width = 7, height = 5)
```

## Bias

```{r}
plot_data[,.(Bias=mean(ests - true_val)), by = .(Setting, A0)] |> knitr::kable()
#plot_data[,.(Bias=mean(ests - true_val)), by = .(Setting, A0)] |> xtable::xtable()
```

## Coverage

```{r}
plot_data[,.(Coverage=mean(Lower <= true_val & true_val <= Upper)), by = .(Setting, A0)] |> knitr::kable()
plot_data[,.(Coverage=mean(Lower <= true_val & true_val <= Upper)), by = .(Setting, A0)] |> xtable::xtable()

```



