---
title: "Recurrent Events with the Random Forest"
output:
  pdf_document:
    toc: true
  html_document:
    toc: true
date: "2026-02-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(randomForestSRC)
library(ggplot2)
library(tidyr)
library(dplyr)
library(gridExtra)
library(simevent)
library(survival)
library(data.table)
library(ggpubr)
```


# The Idea

The idea is, in order to be able to simulate recurrent events data with help of the random forest, to divide data into multiple smaller data sets. A data set for T0 -> T1, T1 -> T2 and so on. For each sub data set, we fit a random forest. We use these and `simEventObj` to simulate recurrent events data. 


# Application

Simulate an original recurrent events data set
```{r}
set.seed(373)
data <- simTreatment(3 * 10^3, cens = 0)
```

We find the sub data set of T0 -> T1, T1 -> T2 and T2 -> T3. We split up into data set depending on whether A = 1 or L = 1 because at_risk changes in between. 
```{r}
# T0 -> T1
data1 <- data[order(Time), .SD[1], by = ID]
setkey(data1, ID)

# T1 -> T2
data2A <- data[ID %in% data1[!(Delta %in% c(0,1)) & A == 1, ID], ]
data2A <- data2A[order(Time), .SD[1:2], by = ID]
data2A[, oldTime := min(Time), by = ID]
data2A[, newTime := Time - oldTime, by = ID]
data2A <- data2A[order(Time), .SD[2], by = ID]
setkey(data2A, ID)

data2L <- data[ID %in% data1[!(Delta %in% c(0,1)) & L == 1, ID], ]
data2L <- data2L[order(Time), .SD[1:2], by = ID]
data2L[, oldTime := min(Time), by = ID]
data2L[, newTime := Time - oldTime, by = ID]
data2L <- data2L[order(Time), .SD[2], by = ID]
setkey(data2L, ID)

# T2 -> T3
data3 <- data[ID %in% c(data2A[!(Delta %in% c(0,1)), ID], data2L[!(Delta %in% c(0,1)), ID]), ]
data3 <- data3[order(Time), .SD[2:3], by = ID]
data3[, oldTime := min(Time), by = ID]
data3[, newTime := Time - oldTime, by = ID]
data3 <- data3[order(Time), .SD[2], by = ID]
setkey(data3, ID)
```

Fit models
```{r}
set.seed(100)
# T0 -> T1
#tune.nodesize(Surv(Time, Delta) ~ L0, data = data1)
RF_fit1 <- rfsrc(Surv(Time, Delta) ~ L0, data = data1, nodesize = 70)

# T1 -> T2
#tune.nodesize(Surv(newTime, Delta) ~ L0, data = data2A)
#tune.nodesize(Surv(newTime, Delta) ~ L0, data = data2L)
RF_fit2A <- rfsrc(Surv(newTime, Delta) ~ L0 + oldTime, data = data2A, nodesize = 60)
RF_fit2L <- rfsrc(Surv(newTime, Delta) ~ L0 + oldTime, data = data2L, nodesize = 20)

# T2 -> T3
#tune.nodesize(Surv(newTime,Delta) ~ L0 + L + A, data = data3)
RF_fit3 <- rfsrc(Surv(newTime, Delta) ~ L0 + oldTime, data = data3, nodesize = 20)
```

Create `predict2`
```{r}
predict2 <- function(obj, ...) {
  UseMethod("predict2")
}
predict2.rfsrc <- function(obj, sim_data, ...){
  preds <- stats::predict(obj, sim_data)
  if(is.na(dim(preds$chf)[3])){
    chf <- array(dim = c(dim(preds$chf), 1))
    chf[,,1] <- preds$chf
  } else {
    chf <- preds$chf
  }
  list(time = preds$time.interest, chf = chf)
}
```

Simulate the recurrent events. First T0 -> T1
```{r}
set.seed(100)
# The old covariates variables
old_vars1 <- as.matrix(data1[, L0])
colnames(old_vars1) <- "L0"
new_data1 <- simEventObj(N = 10^4, 
                         obj = RF_fit1, 
                         old_vars = old_vars1,
                         event_names = c("D", "A", "L"))

# Number of individuals for which we simulate using either RFA or RFL
N_nextA <- nrow(new_data1[A == 1, ])
N_nextL <- nrow(new_data1[L == 1, ])
```

Check simulations
```{r}
new_data1[Delta != 0, table(Delta)]/nrow(new_data1)
data1[, table(Delta)]/nrow(data1)
new_data1[Delta != 0, summary(Time)]
data1[, summary(Time)]

ggarrange(plotEventData(new_data1[Delta != 0][1:1000,], title = "Sim Data"), 
          plotEventData(data1[1:1000,], title = "Original Data"), ncol = 2)
```

Then T1 -> T2
```{r}
# We use the old covariates
old_vars <- data.frame(L0 = new_data1[A == 1, L0], oldTime = new_data1[A == 1, Time])
new_data2A <- simEventObj(N = N_nextA, 
                          obj = RF_fit2A, 
                          event_names = c("D", "L"), 
                          old_vars = old_vars,
                          useOldVars = TRUE)

new_data2A$ID <- new_data1[A == 1, ID]
new_data2A[, Time := Time + oldTime]
# L is coded as Delta = 3 in the original data set
new_data2A[L == 1, Delta := 3] 

old_vars = data.frame(L0 = new_data1[L == 1, L0], oldTime = new_data1[L == 1, Time])
new_data2L <- simEventObj(N = N_nextL, 
                          obj = RF_fit2L, 
                          event_names = c("D", "A"), 
                          old_vars = old_vars,
                          useOldVars = TRUE)
new_data2L$ID <- new_data1[L == 1, ID]
new_data2L[, Time := Time + oldTime]
```

Check simulations
```{r}
new_data2A[Delta != 0, table(Delta)]/nrow(new_data2A)
data2A[, table(Delta)]/nrow(data2A)
new_data2A[Delta != 0, summary(Time)]
data2A[, summary(Time)]

ggarrange(plotEventData(new_data2A[Delta != 0][1:1000,], title = "Sim Data"), 
          plotEventData(data2A[1:1000,], title = "Original Data"), ncol = 2)
```

```{r}
new_data2L[Delta != 0, table(Delta)]/nrow(new_data2L)
data2L[, table(Delta)]/nrow(data2L)
new_data2L[Delta != 0, summary(Time)]
data2L[, summary(Time)]

ggarrange(plotEventData(new_data2L[Delta != 0][1:1000,], title = "Sim Data"), 
          plotEventData(data2L[1:1000,], title = "Original Data"), ncol = 2)
```

Then T2 -> T3
```{r}
N_next <- nrow(new_data2A[!Delta %in% c(0,1)]) + nrow(new_data2L[!Delta %in% c(0,1)])
oldTime <- c(new_data2A[!Delta %in% c(0,1), Time], new_data2L[!Delta %in% c(0,1), Time])   
L0 <- c(new_data2A[!Delta %in% c(0,1), L0], new_data2L[!Delta %in% c(0,1), L0])
old_vars <- data.frame(L0 = L0,
                       oldTime = oldTime) 

new_data3 <- simEventObj(N = N_next, 
                         obj = RF_fit3,
                         event_names = c("D"), 
                         old_vars = old_vars,
                         useOldVars = TRUE)

new_data3$ID <- c(new_data2A[!Delta %in% c(0,1), ID], new_data2L[!Delta%in% c(0,1), ID])
new_data3[, Time := Time + oldTime]
```

```{r}
new_data3[Delta != 0, table(Delta)]/nrow(new_data3)
data3[, table(Delta)]/nrow(data3)
new_data3[Delta != 0, summary(Time)]
data3[, summary(Time)]

ggarrange(plotEventData(new_data3[Delta != 0][1:1000,], title = "Sim Data"), 
          plotEventData(data3[1:1000,], title = "Original Data"), ncol = 2)
```

Collect data into one data set
```{r}
# Correcting Covariates
new_data2A[, A := 1][, oldTime := NULL]
new_data2L[, L := 1][, oldTime := NULL]
new_data3[, c("A", "L") := 1][, oldTime := NULL]

sim_data <- rbind(new_data1, new_data2A, new_data2L, new_data3)
setkey(sim_data, ID, Time)
```

Does the simulated data approximately correspond to the true data
```{r}
sim_data[, table(Delta)]/nrow(sim_data)
data[, table(Delta)]/nrow(data)
sim_data[Delta != 0, summary(Time)]
data[, summary(Time)]

ggarrange(plotEventData(sim_data[Delta != 0][1:2000,], title = "Sim Data"), 
          plotEventData(data[1:2000,], title = "Original Data"), ncol = 2)
```

Sanity check
```{r}
sim_data <- IntFormatData(sim_data, N_cols = 5:7)

survfit1 <- coxph(Surv(tstart, tstop, Delta == 1) ~ L0 + A + L, data = sim_data)
survfit2 <- coxph(Surv(tstart, tstop, Delta == 2) ~ L0 + L, data = sim_data[A == 0])
survfit3 <- coxph(Surv(tstart, tstop, Delta == 3) ~ L0 + A, data = sim_data[L == 0])

# Check estimations
df <- rbind(cbind(c(1, -1, 1),  confint(survfit1), 1),
            cbind(c(1, 1),  confint(survfit2), 2),
            cbind(c(1, -0.5),  confint(survfit3), 3))

colnames(df) <- c("actual", "LowCI", "UpCI", "fit")
df <- data.table(df)
df[, Index := seq_len(.N), by = fit]
df[, In := actual >= LowCI & actual <= UpCI]

ggplot(df, aes(x = actual, y = Index)) +
  geom_segment(aes(x = LowCI, xend = UpCI, yend = Index),
               colour = "grey50", linewidth = 2) +
  geom_point(aes(color = In), size = 2.5) +
  facet_wrap(~fit) +
  theme_bw()+
  scale_color_manual(values = c("FALSE" = "red", "TRUE" = "darkgreen"))
```

# Debugging using cox models

Fit models
```{r}
# T0 -> T1
survfit_death1 <- coxph(Surv(Time, Delta == 1) ~ L0, data = data1)
survfit_oper1 <- coxph(Surv(Time, Delta == 2) ~ L0, data = data1)
survfit_cov1 <- coxph(Surv(Time, Delta == 3) ~ L0, data = data1)

# T1 -> T2
survfit_death2A <- coxph(Surv(newTime, Delta == 1) ~ L0 + oldTime, data = data2A)
survfit_cov2A <- coxph(Surv(newTime, Delta == 3) ~ L0 + oldTime, data = data2A)


survfit_death2L <- coxph(Surv(newTime, Delta == 1) ~ L0 + oldTime, data = data2L)
survfit_oper2L <- coxph(Surv(newTime, Delta == 2) ~ L0 + oldTime, data = data2L)

# T2 -> T3
survfit_death3 <- coxph(Surv(Time, Delta == 1) ~ L0 + oldTime, data = data3)
```

Simulate the recurrent events. First T0 -> T1
```{r}
set.seed(100)
# The old covariates variables
old_vars1 <- as.matrix(data1[, L0, drop = FALSE])
colnames(old_vars1) <- "L0"
cox_fits <- list("D" = survfit_death1, "A" = survfit_oper1, "L" = survfit_cov1)
class(cox_fits) <- "simevent"
new_data1 <- simEventObj(N = 10^4, 
                         obj = cox_fits, 
                         old_vars = old_vars1,
                         event_names = c("D", "A", "L"))

# Number of individuals for which we simulate using either cox fits on A = 1 or Cox fits on L = 1
N_nextA <- nrow(new_data1[A == 1, ])
N_nextL <- nrow(new_data1[L == 1, ])
```

Check simulations
```{r}
new_data1[Delta != 0, table(Delta)]/nrow(new_data1)
data1[, table(Delta)]/nrow(data1)
new_data1[Delta != 0, summary(Time)]
data1[, summary(Time)]

ggarrange(plotEventData(new_data1[Delta != 0][1:2000,], title = "Sim Data"), 
          plotEventData(data1[1:2000,], title = "Original Data"), ncol = 2)
```

Then T1 -> T2
```{r}
# We use the old covariates
old_vars <- data.frame(L0 = new_data1[A == 1, L0], oldTime = new_data1[A == 1, Time])
cox_fits <- list("D" = survfit_death2A, "L" = survfit_cov2A)
class(cox_fits) <- "simevent"
new_data2A <- simEventObj(N = N_nextA, 
                          obj = cox_fits, 
                          event_names = c("D", "L"), 
                          old_vars = old_vars,
                          useOldVars = TRUE)

new_data2A$ID <- new_data1[A == 1, ID]
new_data2A[, Time := Time + oldTime]
# L is coded as Delta = 3 in the original data set
new_data2A[L == 1, Delta := 3] 

old_vars = data.frame(L0 = new_data1[L == 1, L0], oldTime = new_data1[L == 1, Time])
cox_fits <- list("D" = survfit_death2L, "L" = survfit_oper2L)
class(cox_fits) <- "simevent"
new_data2L <- simEventObj(N = N_nextL, 
                          obj = cox_fits, 
                          event_names = c("D", "A"), 
                          old_vars = old_vars,
                          useOldVars = TRUE)
new_data2L$ID <- new_data1[L == 1, ID]
new_data2L[, Time := Time + oldTime]
```

Check simulations
```{r}
new_data2A[Delta != 0, table(Delta)]/nrow(new_data2A)
data2A[, table(Delta)]/nrow(data2A)
new_data2A[Delta != 0, summary(Time)]
data2A[, summary(Time)]

ggarrange(plotEventData(new_data2A[Delta != 0][1:1000,], title = "Sim Data"), 
          plotEventData(data2A[1:1000,], title = "Original Data"), ncol = 2)
```

```{r}
new_data2L[Delta != 0, table(Delta)]/nrow(new_data2L)
data2L[, table(Delta)]/nrow(data2L)
new_data2L[Delta != 0, summary(Time)]
data2L[, summary(Time)]

ggarrange(plotEventData(new_data2L[Delta != 0][1:1000,], title = "Sim Data"), 
          plotEventData(data2L[1:1000,], title = "Original Data"), ncol = 2)
```

Then T2 -> T3
```{r}
N_next <- nrow(new_data2A[!Delta %in% c(0,1)]) + nrow(new_data2L[!Delta %in% c(0,1)])
oldTime <- c(new_data2A[!Delta %in% c(0,1), Time], new_data2L[!Delta %in% c(0,1), Time])   
L0 <- c(new_data2A[!Delta %in% c(0,1), L0], new_data2L[!Delta %in% c(0,1), L0])
old_vars <- data.frame(L0 = L0,
                       oldTime = oldTime) 
cox_fits <- list("D" = survfit_death3)
class(cox_fits) <- "simevent"

new_data3 <- simEventObj(N = N_next, 
                         obj = cox_fits,
                         event_names = c("D"), 
                         old_vars = old_vars,
                         useOldVars = TRUE)

new_data3$ID <- c(new_data2A[!Delta %in% c(0,1), ID], new_data2L[!Delta%in% c(0,1), ID])
new_data3[, Time := Time + oldTime]
```

```{r}
new_data3[Delta != 0, table(Delta)]/nrow(new_data3)
data3[, table(Delta)]/nrow(data3)
new_data3[Delta != 0, summary(Time)]
data3[, summary(Time)]

ggarrange(plotEventData(new_data3[Delta != 0][1:1000,], title = "Sim Data"), 
          plotEventData(data3[1:1000,], title = "Original Data"), ncol = 2)
```

Collect data into one data set
```{r}
# Correcting Covariates
new_data2A[, A := 1][, oldTime := NULL]
new_data2L[, L := 1][, oldTime := NULL]
new_data3[, c("A", "L") := 1][, oldTime := NULL]

sim_data <- rbind(new_data1, new_data2A, new_data2L, new_data3)
setkey(sim_data, ID, Time)
```

Does the simulated data approximately correspond to the true data
```{r}
sim_data[, table(Delta)]/nrow(sim_data)
data[, table(Delta)]/nrow(data)
sim_data[Delta != 0, summary(Time)]
data[, summary(Time)]

ggarrange(plotEventData(sim_data[Delta != 0][1:2000,], title = "Sim Data"), 
          plotEventData(data[1:2000,], title = "Original Data"), ncol = 2)
```

Sanity check
```{r}
sim_data <- IntFormatData(sim_data, N_cols = 5:7)

survfit1 <- coxph(Surv(tstart, tstop, Delta == 1) ~ L0 + A + L, data = sim_data)
survfit2 <- coxph(Surv(tstart, tstop, Delta == 2) ~ L0 + L, data = sim_data[A == 0])
survfit3 <- coxph(Surv(tstart, tstop, Delta == 3) ~ L0 + A, data = sim_data[L == 0])

# Check estimations
df <- rbind(cbind(c(1, -1, 1),  confint(survfit1), 1),
            cbind(c(1, 1),  confint(survfit2), 2),
            cbind(c(1, -0.5),  confint(survfit3), 3))

colnames(df) <- c("actual", "LowCI", "UpCI", "fit")
df <- data.table(df)
df[, Index := seq_len(.N), by = fit]
df[, In := actual >= LowCI & actual <= UpCI]

ggplot(df, aes(x = actual, y = Index)) +
  geom_segment(aes(x = LowCI, xend = UpCI, yend = Index),
               colour = "grey50", linewidth = 2) +
  geom_point(aes(color = In), size = 2.5) +
  facet_wrap(~fit) +
  theme_bw()+
  scale_color_manual(values = c("FALSE" = "red", "TRUE" = "darkgreen"))
```


# Debugging using cox models

Fit models
```{r}
# T0 -> T1
survfit_death1 <- coxph(Surv(Time, Delta == 1) ~ L0, data = data1)
survfit_oper1 <- coxph(Surv(Time, Delta == 2) ~ L0, data = data1)
survfit_cov1 <- coxph(Surv(Time, Delta == 3) ~ L0, data = data1)

# T1 -> T2
survfit_death2A <- coxph(Surv(newTime, Delta == 1) ~ L0 + oldTime, data = data2A)
survfit_cov2A <- coxph(Surv(newTime, Delta == 3) ~ L0 + oldTime, data = data2A)


survfit_death2L <- coxph(Surv(newTime, Delta == 1) ~ L0 + oldTime, data = data2L)
survfit_oper2L <- coxph(Surv(newTime, Delta == 2) ~ L0 + oldTime, data = data2L)

# T2 -> T3
survfit_death3 <- coxph(Surv(Time, Delta == 1) ~ L0 + oldTime, data = data3)
```

Simulate the recurrent events. First T0 -> T1
```{r}
set.seed(100)
# The old covariates variables
old_vars1 <- as.matrix(data1[, L0, drop = FALSE])
colnames(old_vars1) <- "L0"
cox_fits <- list("D" = survfit_death1, "A" = survfit_oper1, "L" = survfit_cov1)
new_data1 <- simEventCox(N = 10^4, 
                         cox_fits = cox_fits, 
                         old_vars = old_vars1,
                         n_event_max = c(1, 1, 1),
                         term_events = c(1, 2, 3))

# Number of individuals for which we simulate using either cox fits on A = 1 or Cox fits on L = 1
N_nextA <- nrow(new_data1[A == 1, ])
N_nextL <- nrow(new_data1[L == 1, ])
```

Check simulations
```{r}
new_data1[, table(Delta)]/nrow(new_data1)
data1[, table(Delta)]/nrow(data1)
new_data1[, summary(Time)]
data1[, summary(Time)]

ggarrange(plotEventData(new_data1[1:2000,], title = "Sim Data"), 
          plotEventData(data1[1:2000,], title = "Original Data"), ncol = 2)
```

Then T1 -> T2
```{r}
# We use the old covariates
old_vars <- data.frame(L0 = new_data1[A == 1, L0], oldTime = new_data1[A == 1, Time])
cox_fits <- list("D" = survfit_death2A, "L" = survfit_cov2A)
new_data2A <- simEventCox(N = N_nextA, 
                          cox_fits = cox_fits, 
                          useOldVars = TRUE,
                          old_vars = old_vars,
                          n_event_max = c(1, 1),
                          term_events = c(1, 2))

new_data2A$ID <- new_data1[A == 1, ID]
new_data2A[, Time := Time + oldTime]
# L is coded as Delta = 3 in the original data set
new_data2A[L == 1, Delta := 3] 
# D is coded as Delta = 1 in the original data set

old_vars = data.frame(L0 = new_data1[L == 1, L0], oldTime = new_data1[L == 1, Time])
cox_fits <- list("D" = survfit_death2L, "L" = survfit_oper2L)
new_data2L <- simEventCox(N = N_nextL, 
                          cox_fits = cox_fits, 
                          useOldVars = TRUE,
                          old_vars = old_vars,
                          n_event_max = c(1, 1),
                          term_events = c(1, 2))

new_data2L$ID <- new_data1[L == 1, ID]
new_data2L[, Time := Time + oldTime]
```

Check simulations
```{r}
new_data2A[Delta != 0, table(Delta)]/nrow(new_data2A)
data2A[, table(Delta)]/nrow(data2A)
new_data2A[Delta != 0, summary(Time)]
data2A[, summary(Time)]

ggarrange(plotEventData(new_data2A[Delta != 0][1:1000,], title = "Sim Data"), 
          plotEventData(data2A[1:1000,], title = "Original Data"), ncol = 2)
```

```{r}
new_data2L[Delta != 0, table(Delta)]/nrow(new_data2L)
data2L[, table(Delta)]/nrow(data2L)
new_data2L[Delta != 0, summary(Time)]
data2L[, summary(Time)]

ggarrange(plotEventData(new_data2L[Delta != 0][1:1000,], title = "Sim Data"), 
          plotEventData(data2L[1:1000,], title = "Original Data"), ncol = 2)
```

Then T2 -> T3
```{r}
N_next <- nrow(new_data2A[!Delta %in% c(0,1)]) + nrow(new_data2L[!Delta %in% c(0,1)])
oldTime <- c(new_data2A[!Delta %in% c(0,1), Time], new_data2L[!Delta %in% c(0,1), Time])   
L0 <- c(new_data2A[!Delta %in% c(0,1), L0], new_data2L[!Delta %in% c(0,1), L0])
old_vars <- data.frame(L0 = L0,
                       oldTime = oldTime) 
cox_fits <- list("D" = survfit_death3)
class(cox_fits) <- "simevent"

new_data3 <- simEventObj(N = N_next, 
                         obj = cox_fits,
                         event_names = c("D"), 
                         old_vars = old_vars,
                         useOldVars = TRUE)

new_data3$ID <- c(new_data2A[!Delta %in% c(0,1), ID], new_data2L[!Delta%in% c(0,1), ID])
new_data3[, Time := Time + oldTime]
```

```{r}
new_data3[Delta != 0, table(Delta)]/nrow(new_data3)
data3[, table(Delta)]/nrow(data3)
new_data3[Delta != 0, summary(Time)]
data3[, summary(Time)]

ggarrange(plotEventData(new_data3[Delta != 0][1:1000,], title = "Sim Data"), 
          plotEventData(data3[1:1000,], title = "Original Data"), ncol = 2)
```

Collect data into one data set
```{r}
# Correcting Covariates
new_data2A[, A := 1][, oldTime := NULL]
new_data2L[, L := 1][, oldTime := NULL]
new_data3[, c("A", "L") := 1][, oldTime := NULL]

sim_data <- rbind(new_data1, new_data2A, new_data2L, new_data3)
setkey(sim_data, ID, Time)
```

Does the simulated data approximately correspond to the true data
```{r}
sim_data[, table(Delta)]/nrow(sim_data)
data[, table(Delta)]/nrow(data)
sim_data[Delta != 0, summary(Time)]
data[, summary(Time)]

ggarrange(plotEventData(sim_data[Delta != 0][1:2000,], title = "Sim Data"), 
          plotEventData(data[1:2000,], title = "Original Data"), ncol = 2)
```

Sanity check
```{r}
sim_data <- IntFormatData(sim_data, N_cols = 5:7)

survfit1 <- coxph(Surv(tstart, tstop, Delta == 1) ~ L0 + A + L, data = sim_data)
survfit2 <- coxph(Surv(tstart, tstop, Delta == 2) ~ L0 + L, data = sim_data[A == 0])
survfit3 <- coxph(Surv(tstart, tstop, Delta == 3) ~ L0 + A, data = sim_data[L == 0])

# Check estimations
df <- rbind(cbind(c(1, -1, 1),  confint(survfit1), 1),
            cbind(c(1, 1),  confint(survfit2), 2),
            cbind(c(1, -0.5),  confint(survfit3), 3))

colnames(df) <- c("actual", "LowCI", "UpCI", "fit")
df <- data.table(df)
df[, Index := seq_len(.N), by = fit]
df[, In := actual >= LowCI & actual <= UpCI]

ggplot(df, aes(x = actual, y = Index)) +
  geom_segment(aes(x = LowCI, xend = UpCI, yend = Index),
               colour = "grey50", linewidth = 2) +
  geom_point(aes(color = In), size = 2.5) +
  facet_wrap(~fit) +
  theme_bw()+
  scale_color_manual(values = c("FALSE" = "red", "TRUE" = "darkgreen"))
```
