---
title: "Recurrent Events with the Random Forest"
output:
  pdf_document:
    toc: true
  html_document:
    toc: true
date: "2026-02-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(randomForestSRC)
library(ggplot2)
library(tidyr)
library(dplyr)
library(gridExtra)
library(simevent)
library(survival)
```


# The Idea

The idea is, in order to be able to simulate recurrent events data with help of the random forest, to divide data into multiple smaller data sets. A data set for T0 -> T1, T1 -> T2 and so on. For each sub data set, we fit a random forest. We use these and `simEventObj` to simulate recurrent events data. 

# Recurrent Event Simulation Function

```{r}
simEventObjRec <- function(N,
                           obj,
                           event_names = NULL,
                           sim_data = NULL,
                           T_old = 0,
                           IDs = 1:N) {
  
  ID <- predict2 <- NULL
  
  # Initialize last event time
  T0 <- rep(0, N)
  
  # The cumulative hazard and inverse cumulative hazard
  y.pred <- predict2(obj, sim_data)                                             # Predictions
  times <-  c(0,y.pred$time)                                                    # Time points
  chf_mat <- y.pred$chf                                                         # Cumulative hazard for simulated data
  num_events <- dim(chf_mat)[3]                                                 # Number of events
  
  # Creating columns for event counts
  if(!is.null(event_names)) for (name in event_names) sim_data[[name]] <- 0 else
    for (name in paste0("N", 1:num_events)) sim_data[[name]] <- 0
  
  # Defining the cumulativ hazard and the inverse cumulative hazard
  cumhaz_fn <- function(t, i, j){
    idx <- findInterval(t, times)
    t1 <- times[idx]; t2 <- times[idx + 1]
    y1 <- chf_mat[cbind(i,idx, j)]; y2 <- chf_mat[cbind(i,(idx + 1), j)]
    y1 + (t - t1) * (y2 - y1) / (t2 - t1)
  }
  
  invcumhaz_fn <- function(p, i, j){
    cumhazz <-  chf_mat[,,j]
    idx <- sapply(1:length(i), FUN = function(k) findInterval(p[k], cumhazz[i[k],]))
    idx[idx == ncol(cumhazz)] <- ncol(cumhazz) - 1
    p1 <- cumhazz[cbind(i, idx)]; p2 <- cumhazz[cbind(i,idx + 1)]
    y1 <- times[idx]; y2 <- times[idx + 1]
    # If the cumulative hazard flattens, we choose the smallest time
    y1 + ifelse(p1 == p2, 0, (p - p1) * (y2 - y1) / (p2 - p1))
  }
  
  # Calculate the cumulative intensity per individual per event
  cum_int_Tk <- matrix(nrow = N, ncol = num_events)
  for(j in seq_len(num_events)) {
    cum_int_Tk[,j] <- cumhaz_fn(T0, 1:N, j)
  }
  
  # Simulate the uniform random variable
  U <- matrix(-log(stats::runif(N * num_events)), ncol = num_events)  # matrix for the random draws
  V <- U + cum_int_Tk
  
  # Find the event times
  event_times <- matrix(nrow = N, ncol = num_events)
  for(j in seq_len(num_events)) {
    event_times[,j] <- invcumhaz_fn(V[,j], 1:N, j)
  }
  
  # The next event is the minimum of these events
  T_k <- apply(event_times, 1, min)
  Deltas <- apply(event_times, 1, which.min)
  
  # Update event counts
  for(i in 1:num_events){
    sim_data[seq_len(N), num_cov + i] <- sim_data[seq_len(N), num_cov + i] + ifelse(Deltas == i, 1, 0)
  }
  
  # Store data
  kth_event <- data.table(ID = IDs,
                          Time = T_k + T_old,
                          Delta = Deltas)
  
  res <- cbind(kth_event, data.table::as.data.table(sim_data))
  data.table::setkey(res, ID)
  
  return(res)
}
```


# Application

Simulate an original recurrent events data set
```{r}
set.seed(373)
data <- simTreatment(1000, cens = 0)
```

We find the sub data set of T0 -> T1, T1 -> T2 and T2 -> T3
```{r}
# T0 -> T1
data1 <- data[order(Time), .SD[1], by = ID]
setkey(data1, ID)

# T1 -> T2
data2A <- data[ID %in% data1[!(Delta %in% c(0,1)) & A == 1, ID], ]
data2A <- data2A[order(Time), .SD[1:2], by = ID]
data2A[, newTime := Time - min(Time), by = ID]
data2A <- data2A[order(Time), .SD[2], by = ID]
setkey(data2A, ID)

data2L <- data[ID %in% data1[!(Delta %in% c(0,1)) & L == 1, ID], ]
data2L <- data2L[order(Time), .SD[1:2], by = ID]
data2L[, newTime := Time - min(Time), by = ID]
data2L <- data2L[order(Time), .SD[2], by = ID]
setkey(data2L, ID)

# T2 -> T3
data3 <- data[ID %in% c(data2A[!(Delta %in% c(0,1)), ID], data2L[!(Delta %in% c(0,1)), ID]), ]
data3 <- data3[order(Time), .SD[2:3], by = ID]
data3[, newTime := Time - min(Time), by = ID]
data3 <- data3[order(Time), .SD[2], by = ID]
setkey(data3, ID)
```

Fit models
```{r}
set.seed(100)
# T0 -> T1
#tune.nodesize(Surv(Time, Delta) ~ L0, data = data1)
RF_fit1 <- rfsrc(Surv(Time, Delta) ~ L0, data = data1, nodesize = 70)

# T1 -> T2
#tune.nodesize(Surv(newTime, Delta) ~ L0, data = data2A)
#tune.nodesize(Surv(newTime, Delta) ~ L0, data = data2L)
RF_fit2A <- rfsrc(Surv(newTime, Delta) ~L0, data = data2A, nodesize = 60)
RF_fit2L <- rfsrc(Surv(newTime, Delta) ~L0, data = data2L, nodesize = 20)

# T2 -> T3
#tune.nodesize(Surv(newTime,Delta) ~ L0 + L + A, data = data3)
RF_fit3 <- rfsrc(Surv(newTime, Delta) ~ L0, data = data3, nodesize = 20)
```

Create `predict2`
```{r}
predict2 <- function(obj, ...) {
  UseMethod("predict2")
}
predict2.rfsrc <- function(obj, sim_data, ...){
  preds <- stats::predict(obj, sim_data)
  if(is.na(dim(preds$chf)[3])){
    chf <- array(dim = c(dim(preds$chf), 1))
    chf[,,1] <- preds$chf
  } else {
    chf <- preds$chf
  }
  list(time = preds$time.interest, chf = chf)
}
```

Simulate the recurrent events
```{r}
set.seed(100)
# T0 -> T1
list_old_vars1 <- list(L0 = data1$L0) 
new_data1 <- simEventObj(N = 10^4, obj = RF_fit1, list_old_vars = list_old_vars1,
                         event_names = c("D", "A", "L"))
N_nextA <- nrow(new_data1[A == 1, ])
N_nextL <- nrow(new_data1[L == 1, ])

# T1 -> T2
new_data2A <- simEventObjRec(N = N_nextA, 
                             obj = RF_fit2A, 
                             event_names = c("D", "L"), 
                             sim_data = data.frame(L0 = new_data1[A == 1, L0]), 
                             T_old = new_data1[A == 1, Time], 
                             IDs = new_data1[A == 1, ID])

new_data2L <- simEventObjRec(N = N_nextL, 
                             obj = RF_fit2L, 
                             event_names = c("D", "A"), 
                             sim_data = data.frame(L0 = new_data1[L == 1, L0]), 
                             T_old = new_data1[L == 1, Time], 
                             IDs = new_data1[L == 1, ID])

# T2 -> T3
N_next <- nrow(new_data2A[Delta != 1]) + nrow(new_data2L[Delta != 1])
sim_data <- data.frame(L0 = c(new_data2A[Delta != 1, L0], new_data2L[Delta != 1, L0])) 
T_old <- c(new_data2A[Delta != 1, Time], new_data2L[Delta != 1, Time])
IDs <- c(new_data2A[Delta != 1, ID], new_data2L[Delta != 1, ID])

new_data3 <- simEventObjRec(N = N_next, 
                            obj = RF_fit3, 
                            event_names = c("D"), 
                            sim_data = sim_data, 
                            T_old = T_old, 
                            IDs = IDs)
```




