---
title: "Recurrent Events with the Random Forest"
output:
  pdf_document:
    toc: true
  html_document:
    toc: true
date: "2026-02-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(randomForestSRC)
library(ggplot2)
library(tidyr)
library(dplyr)
library(gridExtra)
library(simevent)
library(survival)
library(data.table)
library(ggpubr)
```


# The Idea

The idea is, in order to be able to simulate recurrent events data with help of the random forest, to divide data into multiple smaller data sets. A data set for T0 -> T1, T1 -> T2 and so on. For each sub data set, we fit a random forest. We use these and `simEventObj` to simulate recurrent events data. 

# Recurrent Event Simulation Function

```{r}
simEventObjRec <- function(N,
                           obj,
                           event_names = NULL,
                           sim_data = NULL,
                           T_old = 0,
                           IDs = 1:N,
                           Delta = NULL) {
  
  ID <- predict2 <- NULL
  
  # The cumulative hazard and inverse cumulative hazard
  y.pred <- predict2(obj, sim_data)                                             # Predictions
  times <-  c(0,y.pred$time)                                                    # Time points
  chf_mat <- y.pred$chf                                                         # Cumulative hazard for simulated data
  num_events <- dim(chf_mat)[3]                                                 # Number of events
  num_cov <- ncol(sim_data)                                                     # Number of covariates
  if(is.null(Delta)) Delta <- 1:num_events
  
  # Creating columns for event counts
  if(!is.null(event_names)) for (name in event_names) sim_data[[name]] <- 0 else
    for (name in paste0("N", 1:num_events)) sim_data[[name]] <- 0
  
  # Defining the cumulativ hazard and the inverse cumulative hazard
  cumhaz_fn <- function(t, i, j){
    idx <- findInterval(t, times)
    t1 <- times[idx]; t2 <- times[idx + 1]
    y1 <- chf_mat[cbind(i,idx, j)]; y2 <- chf_mat[cbind(i,(idx + 1), j)]
    y1 + (t - t1) * (y2 - y1) / (t2 - t1)
  }
  
  invcumhaz_fn <- function(p, i, j){
    cumhazz <-  chf_mat[,,j]
    idx <- sapply(1:length(i), FUN = function(k) findInterval(p[k], cumhazz[i[k],]))
    idx[idx == 0] <- 1
    idx[idx == ncol(cumhazz)] <- ncol(cumhazz) - 1
    p1 <- cumhazz[cbind(i, idx)]; p2 <- cumhazz[cbind(i,idx + 1)]
    y1 <- times[idx]; y2 <- times[idx + 1]
    # If the cumulative hazard flattens, we choose the smallest time
    y1 + ifelse(p1 == p2, 0, (p - p1) * (y2 - y1) / (p2 - p1))
  }
  
  # Simulate the uniform random variable
  U <- matrix(-log(stats::runif(N * num_events)), ncol = num_events)  # matrix for the random draws
  
  # Find the event times
  event_times <- matrix(nrow = N, ncol = num_events)
  for(j in seq_len(num_events)) {
    event_times[,j] <- invcumhaz_fn(U[,j], 1:N, j)
  }
  
  # The next event is the minimum of these events
  T_k <- apply(event_times, 1, min)
  Deltas <- Delta[apply(event_times, 1, which.min)]
  
  # Update event counts
  for(i in 1:num_events){
    sim_data[seq_len(N), num_cov + i] <- sim_data[seq_len(N), num_cov + i] + ifelse(Deltas == Delta[i], 1, 0)
  }
  
  # Store data
  kth_event <- data.table(ID = IDs,
                          Time = T_k + T_old,
                          Delta = Deltas)
  
  res <- cbind(kth_event, data.table::as.data.table(sim_data))
  data.table::setkey(res, ID)
  
  return(res)
}
```


# Application

Simulate an original recurrent events data set
```{r}
set.seed(373)
data <- simTreatment(2000, cens = 0)
```

We find the sub data set of T0 -> T1, T1 -> T2 and T2 -> T3. We split up into data set depending on whether A = 1 or L = 1 because at_risk changes in between. 
```{r}
# T0 -> T1
data1 <- data[order(Time), .SD[1], by = ID]
setkey(data1, ID)

# T1 -> T2
data2A <- data[ID %in% data1[!(Delta %in% c(0,1)) & A == 1, ID], ]
data2A <- data2A[order(Time), .SD[1:2], by = ID]
data2A[, oldTime := min(Time), by = ID]
data2A[, newTime := Time - oldTime, by = ID]
data2A <- data2A[order(Time), .SD[2], by = ID]
setkey(data2A, ID)

data2L <- data[ID %in% data1[!(Delta %in% c(0,1)) & L == 1, ID], ]
data2L <- data2L[order(Time), .SD[1:2], by = ID]
data2L[, oldTime := min(Time), by = ID]
data2L[, newTime := Time - oldTime, by = ID]
data2L <- data2L[order(Time), .SD[2], by = ID]
setkey(data2L, ID)

# T2 -> T3
data3 <- data[ID %in% c(data2A[!(Delta %in% c(0,1)), ID], data2L[!(Delta %in% c(0,1)), ID]), ]
data3 <- data3[order(Time), .SD[2:3], by = ID]
data3[, oldTime := min(Time), by = ID]
data3[, newTime := Time - oldTime, by = ID]
data3 <- data3[order(Time), .SD[2], by = ID]
setkey(data3, ID)
```

Fit models
```{r}
set.seed(100)
# T0 -> T1
#tune.nodesize(Surv(Time, Delta) ~ L0, data = data1)
RF_fit1 <- rfsrc(Surv(Time, Delta) ~ L0, data = data1, nodesize = 70)

# T1 -> T2
#tune.nodesize(Surv(newTime, Delta) ~ L0, data = data2A)
#tune.nodesize(Surv(newTime, Delta) ~ L0, data = data2L)
RF_fit2A <- rfsrc(Surv(newTime, Delta) ~ L0 + oldTime, data = data2A, nodesize = 60)
RF_fit2L <- rfsrc(Surv(newTime, Delta) ~ L0 + oldTime, data = data2L, nodesize = 20)

# T2 -> T3
#tune.nodesize(Surv(newTime,Delta) ~ L0 + L + A, data = data3)
RF_fit3 <- rfsrc(Surv(newTime, Delta) ~ L0 + oldTime, data = data3, nodesize = 20)
```

Create `predict2`
```{r}
predict2 <- function(obj, ...) {
  UseMethod("predict2")
}
predict2.rfsrc <- function(obj, sim_data, ...){
  preds <- stats::predict(obj, sim_data)
  if(is.na(dim(preds$chf)[3])){
    chf <- array(dim = c(dim(preds$chf), 1))
    chf[,,1] <- preds$chf
  } else {
    chf <- preds$chf
  }
  list(time = preds$time.interest, chf = chf)
}
```

Simulate the recurrent events
```{r}
set.seed(100)
# T0 -> T1
list_old_vars1 <- list(L0 = data1$L0) 
new_data1 <- simEventObj(N = 10^4, obj = RF_fit1, 
                         list_old_vars = list_old_vars1,
                         event_names = c("D", "A", "L"))

N_nextA <- nrow(new_data1[A == 1, ])
N_nextL <- nrow(new_data1[L == 1, ])

# T1 -> T2
new_data2A <- simEventObjRec(N = N_nextA, 
                             obj = RF_fit2A, 
                             event_names = c("D", "L"), 
                             sim_data = data.frame(L0 = new_data1[A == 1, L0], oldTime = new_data1[A == 1, Time]), 
                             T_old = new_data1[A == 1, Time], 
                             IDs = new_data1[A == 1, ID],
                             Delta = c(1,3))

new_data2L <- simEventObjRec(N = N_nextL, 
                             obj = RF_fit2L, 
                             event_names = c("D", "A"), 
                             sim_data = data.frame(L0 = new_data1[L == 1, L0], oldTime = new_data1[L == 1, Time]), 
                             T_old = new_data1[L == 1, Time], 
                             IDs = new_data1[L == 1, ID],
                             Delta = c(1,2))

# T2 -> T3
N_next <- nrow(new_data2A[Delta != 1]) + nrow(new_data2L[Delta != 1])
T_old <- c(new_data2A[Delta != 1, Time], new_data2L[Delta != 1, Time])
sim_data <- data.frame(L0 = c(new_data2A[Delta != 1, L0], new_data2L[Delta != 1, L0]),
                       oldTime = T_old) 

IDs <- c(new_data2A[Delta != 1, ID], new_data2L[Delta != 1, ID])

new_data3 <- simEventObjRec(N = N_next, 
                            obj = RF_fit3, 
                            event_names = c("D"), 
                            sim_data = sim_data, 
                            T_old = T_old, 
                            IDs = IDs)
```

Collect data into one data set

```{r}
# Correcting Covariates
new_data2A[, A := 1][, oldTime := NULL]
new_data2L[, L := 1][, oldTime := NULL]
new_data3[, c("A", "L") := 1][, oldTime := NULL]

sim_data <- rbind(new_data1, new_data2A, new_data2L, new_data3)
setkey(sim_data, ID, Time)

ggarrange(plotEventData(sim_data[1:4000,], title = "Sim Data"), 
          plotEventData(data[1:4000,], title = "Original Data"), ncol = 2)
```
Sanity check
```{r}
sim_data <- IntFormatData(sim_data, N_cols = 5:7)

survfit1 <- coxph(Surv(tstart, tstop, Delta == 1) ~ L0 + A + L, data = sim_data)
survfit2 <- coxph(Surv(tstart, tstop, Delta == 2) ~ L0 + L, data = sim_data[A == 0])
survfit3 <- coxph(Surv(tstart, tstop, Delta == 3) ~ L0 + A, data = sim_data[L == 0])

# Check estimations
df <- rbind(cbind(c(1, -1, 1),  confint(survfit1), 1),
            cbind(c(1, 1),  confint(survfit2), 2),
            cbind(c(1, -0.5),  confint(survfit3), 3))

colnames(df) <- c("actual", "LowCI", "UpCI", "fit")
df <- data.table(df)
df[, Index := seq_len(.N), by = fit]
df[, In := actual >= LowCI & actual <= UpCI]

ggplot(df, aes(x = actual, y = Index)) +
  geom_segment(aes(x = LowCI, xend = UpCI, yend = Index),
               colour = "grey50", linewidth = 2) +
  geom_point(aes(color = In), size = 2.5) +
  facet_wrap(~fit) +
  theme_bw()+
  scale_color_manual(values = c("FALSE" = "red", "TRUE" = "darkgreen"))
```

# Debugging using simEventCox

Fit models
```{r}
# T0 -> T1
survfit_death1 <- coxph(Surv(Time, Delta == 1) ~ L0, data = data1)
survfit_oper1 <- coxph(Surv(Time, Delta == 2) ~ L0, data = data1)
survfit_cov1 <- coxph(Surv(Time, Delta == 3) ~ L0, data = data1)

# T1 -> T2
survfit_death2A <- coxph(Surv(newTime, Delta == 1) ~ L0 + oldTime, data = data2A)
survfit_cov2A <- coxph(Surv(newTime, Delta == 3) ~ L0 + oldTime, data = data2A)


survfit_death2L <- coxph(Surv(newTime, Delta == 1) ~ L0 + oldTime, data = data2L)
survfit_oper2L <- coxph(Surv(newTime, Delta == 2) ~ L0 + oldTime, data = data2L)

# T2 -> T3
survfit_death3 <- coxph(Surv(Time, Delta == 1) ~ L0 + oldTime, data = data3)
```

Simulate the recurrent events.

It is a bit more tricky than seemingly, we need to define a recurrent event simulation function, because we want does exact covariates of our individuals from before. 

```{r}
# Equip with predict2 method
predict2 <- function(obj, ...) {
  UseMethod("predict2")
}
predict2.simevent <- function(obj, sim_data){
  # Base hazards
  basehazz_list <- lapply(obj, function(model) basehaz(model, centered = FALSE))

  # Individual specific term
  cox_term <- lapply(obj, function(model)
    exp(stats::predict(model, newdata = sim_data, type="lp", reference = "zero")))

  # Chf
  chf_list <- lapply(1:length(obj), function(j) cox_term[[j]] %*% t(basehazz_list[[j]][["hazard"]]))
  chf <- array(dim = c(c(dim(chf_list[[1]])), length(obj)))
  for(j in 1:length(obj)) chf[,,j] <- chf_list[[j]]

  # Return list
  list(time = basehazz_list[[1]][["time"]], chf = chf)
}
```

```{r}
set.seed(100)
# T0 -> T1
list_old_vars1 <- list(L0 = data1$L0) 
cox_fits1 <- list("D" = survfit_death1, "A" = survfit_oper1,"L" = survfit_cov1)
class(cox_fits1) <- "simevent"
new_data1 <- simEventObj(N = 5*10^4, 
                         obj = cox_fits1,
                         list_old_vars = list_old_vars1,
                         event_names = c("D", "A", "L"))

N_nextA <- nrow(new_data1[A == 1, ])
N_nextL <- nrow(new_data1[L == 1, ])

# T1 -> T2
cox_fits2A <- list("D" = survfit_death2A, "L" = survfit_cov2A)
class(cox_fits2A) <- "simevent"
new_data2A <- simEventObjRec(N = N_nextA, 
                             obj = cox_fits2A, 
                             event_names = c("D", "L"), 
                             sim_data = data.frame(L0 = new_data1[A == 1, L0], oldTime = new_data1[A == 1, Time]), 
                             T_old = new_data1[A == 1, Time], 
                             IDs = new_data1[A == 1, ID],
                             Delta = c(1,3))

cox_fits2L <- list("D" = survfit_death2L, "A" = survfit_oper2L)
class(cox_fits2L) <- "simevent"
new_data2L <- simEventObjRec(N = N_nextL, 
                             obj = cox_fits2L, 
                             event_names = c("D", "A"), 
                             sim_data = data.frame(L0 = new_data1[L == 1, L0], oldTime = new_data1[L == 1, Time]), 
                             T_old = new_data1[L == 1, Time], 
                             IDs = new_data1[L == 1, ID])

# T2 -> T3
N_next <- nrow(new_data2A[Delta != 1]) + nrow(new_data2L[Delta != 1])
T_old <- c(new_data2A[Delta != 1, Time], new_data2L[Delta != 1, Time])
sim_data <- data.frame(L0 = c(new_data2A[Delta != 1, L0], new_data2L[Delta != 1, L0]),
                       oldTime= T_old) 
IDs <- c(new_data2A[Delta != 1, ID], new_data2L[Delta != 1, ID])
cox_fits3 <- list("D" = survfit_death3)
class(cox_fits3) <- "simevent"

new_data3 <- simEventObjRec(N = N_next, 
                            obj = cox_fits3, 
                            event_names = c("D"), 
                            sim_data = sim_data, 
                            T_old = T_old, 
                            IDs = IDs)
```

Collect data into one data set

```{r}
# Correcting Covariates
new_data2A[, A := 1][, oldTime := NULL]
new_data2L[, L := 1][, oldTime := NULL]
new_data3[, c("A", "L") := 1][, oldTime := NULL]
sim_data <- rbind(new_data1, new_data2A, new_data2L, new_data3)
setkey(sim_data, ID, Time)
```


Checking to see whether the simulation approach works
```{r}
sim_data[, table(Delta)]/nrow(sim_data)
data[, table(Delta)]/nrow(data)
sim_data[, summary(Time)]
data[, summary(Time)]
```



```{r}
ggarrange(plotEventData(sim_data[1:2000,], title = "Sim Data"), 
          plotEventData(data[1:2000,], title = "Original Data"), ncol = 2)
```
We still get some really large times...


Debugging: check that the simulation functions work. Start with the cox and then simeventobj and simple scenarios. 

