---
title: "RF Predicitions"
output:
  pdf_document: default
  html_document: default
date: "2025-11-13"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(randomForestSRC)
library(ggplot2)
library(tidyr)
library(dplyr)
library(gridExtra)
library(simevent)
library(survival)
```


# Fitting Random Forest Models

## On Data simulated from Cox

### Survival 

Simulating data, fitting model, predicting, 10 times. 
```{r}
set.seed(375)
beta = matrix(c(0, 0,0.5,-0.5), ncol = 2, nrow = 2)
newdata <- data.frame(L0 = c(0.5,0.5), A0 = c(1,0))
y.pred <- list()

for (i in 1:10){
  data <- simSurvData(N = 2000, beta = beta)
  obj <- rfsrc(Surv(Time, Delta == 1) ~ L0 + A0, data = data, nodesize = 10)
   
  y.pred[[i]] <- predict(obj, newdata = newdata)
}
```

True survival function:
```{r} 
true_surv_death_func1 <- function(t) exp( - 0.1  * t^(1.1) * exp(sum(newdata[1,] * c(0.5,-0.5))))
true_surv_death_func2 <- function(t) exp( - 0.1  * t^(1.1) * exp(sum(newdata[2,] * c(0.5,-0.5))))
# Evaluate the true survival functions at these time points
time_points <- round(y.pred[[1]]$time.interest, 2)
true_surv1 <- true_surv_death_func1(time_points)
true_surv2 <- true_surv_death_func2(time_points)
```

Visualizing fit
```{r} 
par(mfrow = c(1,2), cex.axis = 2.0, cex.lab = 2.0, cex.main = 2.0, mar = c(6,6,1,1), mgp = c(4,1,0))

# Initialize plot for process 1
plot(NULL, xlim = range(y.pred[[1]]$time.interest), ylim = c(0,1),
     xlab = "Time (Year)", ylab = "Survival")

# Loop over all elements in y.pred
for (i in seq_along(y.pred)) {
  lines(y.pred[[i]]$time.interest, y.pred[[i]]$survival[1,],
        col = rgb(0,0,1,alpha=0.3))  # blue semi-transparent
}
lines(time_points, true_surv1, col="blue", lty=2, lwd=2)

# Initialize plot for process 2
plot(NULL, xlim = range(y.pred[[1]]$time.interest), ylim = c(0,1),
     xlab = "Time (Year)", ylab = "Survival")

# Loop over all elements in y.pred
for (i in seq_along(y.pred)) {
  lines(y.pred[[i]]$time.interest, y.pred[[i]]$survival[2,],
        col = rgb(1,0,0,alpha=0.3))  # red semi-transparent
}
lines(time_points, true_surv2, col="red", lty=2, lwd=2)

legend("topright",
       legend=c("A0=1 (Predicted)", "A0=0 (Predicted)", "A0=1 (True)", "A0=0 (True)"),
       col=c("blue","red","blue","red"),
       lty=c(1,1,2,2), lwd=2, cex=1)
```


### Competing Risk

Simulate data, fit RSF model, create predictions
```{r}
set.seed(373)
beta = matrix(c(0.5,-1,-0.5,0.5,0,0.5), ncol = 3, nrow = 2)
newdata <- data.frame(L0 = c(0.5,0.5), A0 = c(1,0))
y.pred <- list()
for(i in 1:10){
  data <- simCRdata(N = 3000, beta = beta)
  
  #tune.nodesize(Surv(Time,Delta) ~ L0 + A0, data = data)
  RF <- rfsrc(Surv(Time, Delta) ~ L0 + A0, data = data, nodesize = 150)

  y.pred[[i]] <- predict(RF, newdata = newdata)
}
```

True CHF
```{r}
true_chf_func1 <- function(t, cov)  0.1  * t^(1.1) * exp(sum(cov * c(-0.5,0.5)))
true_chf_func2 <- function(t, cov)  0.1  * t^(1.1) * exp(sum(cov * c(0,0.5)))

# Evaluate the true survival functions at time points
time_points <- y.pred[[length(y.pred)]]$time.interest
true_chf11 <- true_chf_func1(time_points, newdata[1,])
true_chf12 <- true_chf_func1(time_points, newdata[2,])
true_chf21 <- true_chf_func2(time_points, newdata[1,])
true_chf22 <- true_chf_func2(time_points, newdata[2,])
```

Plot the true CHF together with the 10 estimated CHF´s
```{r}
par(mfrow = c(2, 2), mar   = c(2, 2, 2, 1), cex.axis = 1, 
    cex.lab  = 1, cex.main = 1, mgp = c(2.5, 0.8, 0))  #oma   = c(5, 5, 2, 1)

# ---- Process 1, A0 = 1 ----
plot(time_points, true_chf11, type="l",
     xlab="Time", ylab="CHF",
     col=1, lty=1, lwd=2, xlim=c(0,10), ylim=c(0, 2.5))
for (i in seq_along(y.pred)) {
  lines(y.pred[[i]]$time.interest, y.pred[[i]]$chf[1,,1],
        col=adjustcolor(1, alpha.f=0.3), lty=2, lwd=2)
}
title(main = "Process 1, A0 = 1")

# ---- Process 1, A0 = 0 ----
plot(time_points, true_chf12, type="l",
     xlab="Time", ylab="CHF",
     col=2, lty=1, lwd=2, xlim=c(0,10), ylim=c(0, 1))
for (i in seq_along(y.pred)) {
  lines(y.pred[[i]]$time.interest, y.pred[[i]]$chf[2,,1],
        col=adjustcolor(2, alpha.f=0.3), lty=2, lwd=2)
}
title(main = "Process 1, A0 = 0")

# ---- Process 2, A0 = 1 ----
plot(time_points, true_chf21, type="l",
     xlab="Time", ylab="CHF",
     col=3, lty=1, lwd=2, xlim=c(0,10), ylim=c(0, 2.5))
for (i in seq_along(y.pred)) {
  lines(y.pred[[i]]$time.interest, y.pred[[i]]$chf[1,,2],
        col=adjustcolor(3, alpha.f=0.3), lty=2, lwd=2)
}
title(main = "Process 2, A0 = 1")

# ---- Process 2, A0 = 0 ----
plot(time_points, true_chf22, type="l",
     xlab="Time", ylab="CHF",
     col=4, lty=1, lwd=2, xlim=c(0,10), ylim=c(0, 2))
for (i in seq_along(y.pred)) {
  lines(y.pred[[i]]$time.interest, y.pred[[i]]$chf[2,,2],
        col=adjustcolor(4, alpha.f=0.3), lty=2, lwd=2)
}
title(main = "Process 2, A0 = 0")
```

Fit Cox models
```{r}
cox1 <- coxph(Surv(Time, Delta == 1) ~ L0 + A0, data = data)
cox2 <- coxph(Surv(Time, Delta == 2) ~ L0 + A0, data = data)
basehazz1 <- basehaz(cox1, centered = TRUE)
basehazz2 <- basehaz(cox2, centered = TRUE)
cox_term1 <- exp(predict(cox1, newdata=newdata, type="lp"))
cox_term2 <- exp(predict(cox2, newdata=newdata, type="lp"))
cum_int1 <- outer(basehazz1[['hazard']], cox_term1, "*")
cum_int2 <- outer(basehazz2[['hazard']], cox_term2, "*")
```

Plot a single fit with Cox models:
```{r}
par(mfrow = c(1,2), cex.axis = 1.0, cex.lab = 1.0, cex.main = 2.0, mar = c(3.0,3,1,1), mgp = c(4, 1, 0))
# Process 1
plot(time_points,y.pred[[length(y.pred)]]$chf[1,,1], type="l", xlab="Time",
     ylab="Process 1 CHF", col=1, lty=1, lwd=2, xlim = c(0,15))
lines(time_points, y.pred[[length(y.pred)]]$chf[2,,1], col=1, lty=2, lwd=2)
# Add lines for the true survival functions
lines(time_points, true_chf11, col=2, lty=1, lwd=2)  
lines(time_points, true_chf12, col=2, lty=2, lwd=2)  
lines(basehazz1$time, cum_int1[,1], col=3, lty=1, lwd=2)
lines(basehazz2$time, cum_int1[,2], col=3, lty=2, lwd=2)

# Process 2
plot(time_points,y.pred[[length(y.pred)]]$chf[1,,2], type="l", xlab="Time",
     ylab="Process 2 CHF", col=1, lty=1, lwd=2, xlim = c(0,15))
lines(time_points, y.pred[[length(y.pred)]]$chf[2,,2], col=1, lty=2, lwd=2)
# Add lines for the true survival functions
lines(time_points, true_chf21, col=2, lty=1, lwd=2)  
lines(time_points, true_chf22, col=2, lty=2, lwd=2) 
lines(basehazz2$time, cum_int2[,1], col=3, lty=1, lwd=2)
lines(basehazz2$time, cum_int2[,2], col=3, lty=2, lwd=2)

legend("topright", 
       legend = c("RF CHF A0 = 1", "RF CHF A0 = 0", "True CHF A0 = 1", "True CHF A0 = 0",
                  "Cox CHF A0 = 1", "Cox CHF A0 = 0"),
       col = c(1, 1, 2, 2, 3, 3),
       lty = c(1, 2, 1, 2, 1, 2),
       lwd = 2,
       cex = 0.75)
```


## On Data not from Cox

### Survival 

```{r}
beta <- matrix(c(0, 0, 0, 0, -0.5 , 1, 0, 0), ncol = 2)
beta_TV <- matrix(c(0, 0, 0, 0, 0.5 , -2, 0, 0), ncol = 2)
t_prime <- 2
data <- simEventTV(10^4, beta = beta, term_deltas = c(0,1), tv_eff = beta_TV, t_prime = t_prime)
#Fit model
obj <- rfsrc(Surv(Time, Delta) ~ L0 + A0, data = data, nodesize = 10)
# Predictions
newdata <- data.frame(L0 = c(0.5,0.5), A0 = c(1,0))
y.pred <- predict(obj, newdata = newdata)
# True survival function:
true_surv_death_func1 <- function(t) {
  if(t <= t_prime) return(exp( - 0.1  * t^(1.1) * exp(sum(newdata[1,] * c(-0.5,1)))))
  exp( - (0.1  * t_prime^(1.1) * exp(sum(newdata[1,] * c(-0.5,1))) + 
          0.1  * t^(1.1) * exp(sum(newdata[1,] * c(0,-1))) -
          0.1  * t_prime^(1.1) * exp(sum(newdata[1,] * c(0,-1)))))
}
true_surv_death_func2 <- function(t) {
  if(t <= t_prime) return(exp( - 0.1  * t^(1.1) * exp(sum(newdata[2,] * c(-0.5,1)))))
  exp( - (0.1  * t_prime^(1.1) * exp(sum(newdata[2,] * c(-0.5,1))) + 
          0.1  * t^(1.1) * exp(sum(newdata[2,] * c(0,-1))) -
          0.1  * t_prime^(1.1) * exp(sum(newdata[2,] * c(0,-1)))))
}
# Evaluate the true survival functions at these time points
time_points <- y.pred$time.interest
true_surv1 <- sapply(time_points, FUN = true_surv_death_func1)
true_surv2 <- sapply(time_points, FUN = true_surv_death_func2)
```

Visualizing fit
```{r} 
par(mfrow = c(1,1), cex.axis = 2.0, cex.lab = 2.0, cex.main = 2.0, mar = c(6.0,6,1,1), mgp = c(4, 1, 0))
plot(round(y.pred$time.interest,2), y.pred$survival[1,], type="l", xlab="Time (Year)",
     ylab="Survival", col=1, lty=1, lwd=2, xlim = c(0,5))
lines(round(y.pred$time.interest,2), y.pred$survival[2,], col=2, lty=1, lwd=2)
# Add lines for the true survival functions
lines(time_points, true_surv1, col=1, lty=2, lwd=2) 
lines(time_points, true_surv2, col=2, lty=2, lwd=2)  
legend("topright", 
       legend=c("A0=1 (Predicted)", "A0=0 (Predicted)", "A0=1 (True)", "A0=0 (True)"), 
       col=c(1, 2, 1, 2), 
       lty=c(1, 1, 2, 2), 
       cex=0.75, 
       lwd=2)
```

### Competing Risk

```{r}
beta <- matrix(c(0, 0, 0, 0, 0, -0.5 , 0.5, 0, 0, 0, 0.5, -0.5, 0, 0, 0), ncol = 3)
beta_TV <- matrix(c(0, 0, 0, 0, 0, 1 , -1.5, 0, 0, 0, -1, 1, 0, 0, 0), ncol = 3)
t_prime <- 2
data <- simEventTV(3000, beta = beta, term_deltas = c(0,1,2), tv_eff = beta_TV, t_prime = t_prime)
```

Fit RSF model
```{r}
#tune.nodesize(Surv(Time,Delta) ~ L0 + A0, data = data)
RF <- rfsrc(Surv(Time, Delta) ~ L0 + A0, data = data, nodesize = 150)
```

New Data
```{r}
newdata <- data.frame(L0 = c(0.5,0.5), A0 = c(1,0))
y.pred <- predict(RF, newdata = newdata)
```

True CHF
```{r}
true_chf_func1 <- function(t, cov) {
  if(t <= t_prime) return(0.1  * t^(1.1) * exp(sum(cov * c(-0.5,0.5))))
  0.1  * t_prime^(1.1) * exp(sum(cov * c(-0.5,0.5))) + 
  0.1  * t^(1.1) * exp(sum(cov * c(0.5,-1))) -
  0.1  * t_prime^(1.1) * exp(sum(cov * c(0.5,-1)))
}
true_chf_func2 <- function(t, cov) {
  if(t <= t_prime) return(0.1  * t^(1.1) * exp(sum(cov * c(0.5,-0.5))))
  0.1  * t_prime^(1.1) * exp(sum(cov * c(0.5,-0.5))) + 
  0.1  * t^(1.1) * exp(sum(cov * c(-0.5,0.5))) -
  0.1  * t_prime^(1.1) * exp(sum(cov * c(-0.5,0.5)))
}

# Evaluate the true CIF functions at time points
time_points <- y.pred$time.interest
true_chf11 <- sapply(time_points, FUN = function(t) true_chf_func1(t, newdata[1,]))
true_chf12 <- sapply(time_points, FUN = function(t) true_chf_func1(t, newdata[2,]))
true_chf21 <- sapply(time_points, FUN = function(t) true_chf_func2(t, newdata[1,]))
true_chf22 <- sapply(time_points, FUN = function(t) true_chf_func2(t, newdata[2,]))
```

Fit Cox models
```{r}
cox1 <- coxph(Surv(Time, Delta == 1) ~ L0 + A0, data = data)
cox2 <- coxph(Surv(Time, Delta == 2) ~ L0 + A0, data = data)
basehazz1 <- basehaz(cox1, centered = TRUE)
basehazz2 <- basehaz(cox2, centered = TRUE)
cox_term1 <- exp(predict(cox1, newdata=newdata, type="lp"))
cox_term2 <- exp(predict(cox2, newdata=newdata, type="lp"))
cum_int1 <- outer(basehazz1[['hazard']], cox_term1, "*")
cum_int2 <- outer(basehazz2[['hazard']], cox_term2, "*")
```

Plot:
```{r}
par(mfrow = c(1,2), cex.axis = 2.0, cex.lab = 2.0, cex.main = 2.0, mar = c(6.0,6,1,1), mgp = c(4, 1, 0))
# Process 1
plot(time_points,y.pred$chf[1,,1], type="l", xlab="Time",
     ylab="Process 1 CHF", col=1, lty=1, lwd=2, xlim = c(0,5), ylim = c(0,75))
lines(time_points, y.pred$chf[2,,1], col=1, lty=2, lwd=2)
# Add lines for the true survival functions
lines(time_points, true_chf11, col=2, lty=1, lwd=2)  
lines(time_points, true_chf12, col=2, lty=2, lwd=2)  
lines(basehazz1$time, cum_int1[,1], col=3, lty=1, lwd=2)
lines(basehazz2$time, cum_int1[,2], col=3, lty=2, lwd=2)

# Process 2
plot(time_points,y.pred$chf[1,,2], type="l", xlab="Time",
     ylab="Process 2 CHF", col=1, lty=1, lwd=2, xlim = c(0,5), ylim = c(0,75))
lines(time_points, y.pred$chf[2,,2], col=1, lty=2, lwd=2)
# Add lines for the true survival functions
lines(time_points, true_chf21, col=2, lty=1, lwd=2)  
lines(time_points, true_chf22, col=2, lty=2, lwd=2) 
lines(basehazz2$time, cum_int2[,1], col=3, lty=1, lwd=2)
lines(basehazz2$time, cum_int2[,2], col=3, lty=2, lwd=2)

legend("topright", 
       legend = c("RF CHF A0 = 1", "RF CHF A0 = 0", "True CHF A0 = 1", "True CHF A0 = 0",
                  "Cox CHF A0 = 1", "Cox CHF A0 = 0"),
       col = c(1, 1, 2, 2, 3, 3),
       lty = c(1, 2, 1, 2, 1, 2),
       lwd = 2,
       cex = 0.75)
```

### General Event History

Yet to be implemented...
