  ---
title: "sim_study_RF_vs_Cox"
output: html_document
date: "2025-12-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(randomForestSRC)
library(ggplot2)
library(tidyr)
library(dplyr)
library(gridExtra)
library(simevent)
library(survival)
library(data.table)
library(purrr)
```


# On `follic` Data

We use the `follic` data from the `randomForestSRC` package. 
```{r}
data(follic, package = "randomForestSRC")
```

We fit a model.  
```{r}
set.seed(865)
#tune.nodesize(Surv(time, status) ~ ., follic)
RF_fit <- rfsrc(Surv(time, status) ~ age + hgb + clinstg + ch, follic, nodesize = 20, nsplit = 3)
```

## Population absolute risk

Simulate a new very large data set. This will be our truth.
```{r}
list_old_vars <- list(age = follic$age, hgb = follic$hgb, clinstg = follic$clinstg, 
                      ch = follic$ch)
true_data <- simEventRF(N = 10^6, RF_fit, list_old_vars = list_old_vars, term_events = c(1,2))
```

We calculate the "true" absolute risk function based on the large data set:
```{r}
tis <- seq(0.1,25, length.out = 200)
CIF_true1 <- numeric(200)
CIF_true2 <- numeric(200)
i <- 1

for(ti in tis){
  CIF_true1[i] <- mean(true_data[,(Time <= ti) & Delta == 1])
  CIF_true2[i] <- mean(true_data[,(Time <= ti) & Delta == 2])
  i <- i + 1 
}
```

```{r, warning = FALSE}
ggplot(tibble(time = tis, true_vals1 = CIF_true1, true_vals2 = CIF_true2))+
  geom_line(aes(x = time, y = true_vals1, color = "P1 True CIF"), linewidth = 1.1)+
  geom_line(aes(x = time, y = true_vals2, color = "P2 True CIF"), linewidth = 1.1)+
  scale_color_manual(values = c("P1 True CIF" = "hotpink", "P2 True CIF" = "forestgreen"))+
  labs(title = "Absolute Risk", y = "Absolute Risk")+
  theme_bw()+
  xlim(c(0,24))+
  ylim(c(0,0.6))

ggsave("TruePopAbsRisk.pdf")
```

We then simulate 100 normal sized data sets using `simEventRF`. For each new data set we fit a random forest and cox proportional hazard models. 

```{r}
N <- 600
new_data <- list(); RF_fits <- list(); Cox_fits1 <- list(); Cox_fits2 <- list();
for(i in 1:100){
  # Simulate normal size data set
  new_data[[i]] <- simEventRF(N = N, RF_fit, list_old_vars = list_old_vars, term_events = c(1,2))
  # RF fit
  RF_fits[[i]] <- rfsrc(Surv(Time, Delta) ~ age + hgb + clinstg + ch, 
                        new_data[[i]], nsplit = 3, ntree = 100)
  # A Cox fit for process 1
  Cox_fits1[[i]] <- coxph(Surv(Time, Delta == 1) ~  age + hgb + clinstg + ch, 
                         data = new_data[[i]])
  # A Cox fit for process 2
  Cox_fits2[[i]] <- coxph(Surv(Time, Delta == 2) ~  age + hgb + clinstg + ch, 
                         data = new_data[[i]])
}
```

We calculate the population absolute risk using the fitted objects

```{r}
CIF_vals1 <- list()
CIF_vals2 <- list()
timesCox <- list()
timesRF <- list()
AR1 <- list()
AR2 <- list()

for(i in 1:100){
  # RF
  timesRF[[i]] <- RF_fits[[i]]$time.interest
  CIF_vals1[[i]] <- RF_fits[[i]]$cif[,,1] |> colMeans()
  CIF_vals2[[i]] <- RF_fits[[i]]$cif[,,2] |> colMeans()
  
  # Cox
  # Linear predictors
  exp_terms1 <- exp(predict(Cox_fits1[[i]], type="lp", reference = "zero"))
  exp_terms2 <- exp(predict(Cox_fits2[[i]], type="lp", reference = "zero"))
  # Times
  timesCox[[i]] <- basehaz(Cox_fits1[[i]])$time
  # Cause specific cumulative hazards
  Haz1 <- exp_terms1 %*% t(basehaz(Cox_fits1[[i]], centered = FALSE)$hazard)
  Haz2 <- exp_terms2 %*% t(basehaz(Cox_fits2[[i]], centered = FALSE)$hazard)
  # Survival function
  S <- exp(-(Haz1 + Haz2))
  # Absolute risk calculation
  dLambda1 <- t(apply(Haz1, 1, function(x) c(x[1], diff(x))))
  dLambda2 <- t(apply(Haz2, 1, function(x) c(x[1], diff(x))))
  AR1[[i]] <- t(apply(S * dLambda1, 1, cumsum))
  AR1[[i]] <- colMeans(AR1[[i]])
  AR2[[i]] <- t(apply(S * dLambda2, 1, cumsum))
  AR2[[i]] <- colMeans(AR2[[i]])
}
```

Creating a plot data.frame

```{r}
df_plot <- map_dfr(1:100, function(i) {
  tibble(
    time  = c(timesRF[[i]],
              timesRF[[i]],
              timesCox[[i]],
              timesCox[[i]]),
    value = c(CIF_vals1[[i]],
              CIF_vals2[[i]],
              AR1[[i]],
              AR2[[i]]),
    type  = c(rep("RF CIF", 2*length(timesRF[[i]])),
              rep("Cox CIF", 2*length(timesCox[[i]]))),
    process = c(rep("P1", length(timesRF[[i]])),
              rep("P2", length(timesRF[[i]])),
              rep("P1", length(timesCox[[i]])),
              rep("P2", length(timesCox[[i]]))),
    run = i
  )
})
```

Calculating mean in timepoints

```{r}
df_mean_cox1 <- matrix(nrow = 100, ncol = length(tis))
df_mean_cox2 <- matrix(nrow = 100, ncol = length(tis))
df_mean_RF1 <- matrix(nrow = 100, ncol = length(tis))
df_mean_RF2 <- matrix(nrow = 100, ncol = length(tis))

for(i in 1:100){
  df_mean_cox2[i,] <- approx(x = timesCox[[i]], y = AR2[[i]], xout = tis)$y
  df_mean_cox1[i,] <- approx(x = timesCox[[i]], y = AR1[[i]], xout = tis)$y
  df_mean_RF1[i,] <- approx(x = timesRF[[i]], y = CIF_vals1[[i]], xout = tis)$y
  df_mean_RF2[i,] <- approx(x = timesRF[[i]], y = CIF_vals2[[i]], xout = tis)$y
}

df_mean_cox1 <- colMeans(df_mean_cox1)
df_mean_cox2 <- colMeans(df_mean_cox2)
df_mean_RF1 <- colMeans(df_mean_RF1)
df_mean_RF2 <-colMeans(df_mean_RF2)

df_plot0 <- tibble(
  time  = rep(tis, 6),                                             
  value = c(CIF_true1, CIF_true2, df_mean_cox1, df_mean_cox2, df_mean_RF1, df_mean_RF2),  
  type  = c(rep("True CIF", 2*length(tis)),                                           
            rep("Mean Cox CIF", 2*length(tis)),
            rep("Mean RF CIF", 2*length(tis))),
  process = rep(c(rep("P1", length(tis)), rep("P2", length(tis))), 3),
  run = 0)

df_plot <- rbind(df_plot, df_plot0)
df_plot1 <- df_plot[df_plot$process == "P1", ]
df_plot2 <- df_plot[df_plot$process == "P2", ]
```

```{r}
ggplot() +

  ## 1) simulation runs
  geom_line(
    data = df_plot1 %>% filter(type %in% c("RF CIF", "Cox CIF")),
    aes(time, value,
        color = type,
        linetype = type,
        group = interaction(type, run)),
    alpha = 0.4,
    linewidth = 0.9
  ) +

  ## 2) true + mean curves
  geom_line(
    data = df_plot1 %>% filter(type %in% c("True CIF", "Mean Cox CIF", "Mean RF CIF")),
    aes(time, value,
        color = type,
        linetype = type),
    linewidth = 1.2
  ) +

  scale_color_manual(values = c(
    "True CIF"      = "darkgreen",
    "RF CIF"        = "pink",
    "Cox CIF"       = "lightblue",
    "Mean Cox CIF"  = "blue4",
    "Mean RF CIF"   = "hotpink"
  )) +

  scale_linetype_manual(values = c(
    "True CIF"      = 1,
    "RF CIF"        = 2,
    "Cox CIF"       = 2,
    "Mean Cox CIF"  = 1,
    "Mean RF CIF"   = 1
  )) +

  labs(title = "Absolute Risk (P1)", y = "Abs Risk") +
  theme_bw() +
  xlim(0, 24) +
  ylim(0, 0.6)

ggsave("estimation1.pdf", width = 9, height = 6)
```



```{r, warning = FALSE}
ggplot() +

  ## 1) simulation runs
  geom_line(
    data = df_plot2 %>% filter(type %in% c("RF CIF", "Cox CIF")),
    aes(time, value,
        color = type,
        linetype = type,
        group = interaction(type, run)),
    alpha = 0.4,
    linewidth = 0.9
  ) +

  ## 2) true + mean curves
  geom_line(
    data = df_plot2 %>% filter(type %in% c("True CIF", "Mean Cox CIF", "Mean RF CIF")),
    aes(time, value,
        color = type,
        linetype = type),
    linewidth = 1.2
  ) +

  scale_color_manual(values = c(
    "True CIF"      = "darkgreen",
    "RF CIF"        = "pink",
    "Cox CIF"       = "lightblue",
    "Mean Cox CIF"  = "blue4",
    "Mean RF CIF"   = "hotpink"
  )) +

  scale_linetype_manual(values = c(
    "True CIF"      = 1,
    "RF CIF"        = 2,
    "Cox CIF"       = 2,
    "Mean Cox CIF"  = 1,
    "Mean RF CIF"   = 1
  )) +

  labs(title = "Absolute Risk (P2)", y = "Abs Risk") +
  theme_bw() +
  xlim(0, 24) +
  ylim(0, 0.3)

ggsave("estimation2.pdf", width = 9, height = 6)
```

## Absolute risk stratified by chemo therapy

We create two new large datasets. One where everybody is on therapy, and one where nobody is on therapy.  

```{r}
list_old_vars0 <- list(age = follic$age, hgb = follic$hgb, clinstg = follic$clinstg, 
                      ch =  factor(rep("N", length(follic$ch))))
list_old_vars1 <- list(age = follic$age, hgb = follic$hgb, clinstg = follic$clinstg, 
                      ch = factor(rep("Y", length(follic$ch))))

true_data0 <- simEventRF(N = 10^6, RF_fit, list_old_vars = list_old_vars0, term_events = c(1,2))
true_data1 <- simEventRF(N = 10^6, RF_fit, list_old_vars = list_old_vars1, term_events = c(1,2))
```

We calculate the "true" absolute risk function of the individuals on treatment and the individuals not on treatment based on the large data set:

```{r}
tis <- seq(0.1,25, length.out = 200)
CIF0_true1 <- numeric(200)
CIF0_true2 <- numeric(200)
CIF1_true1 <- numeric(200)
CIF1_true2 <- numeric(200)
i <- 1

for(ti in tis){
  CIF0_true1[i] <- mean(true_data0[,(Time <= ti) & Delta == 1])
  CIF0_true2[i] <- mean(true_data0[,(Time <= ti) & Delta == 2])
  CIF1_true1[i] <- mean(true_data1[,(Time <= ti) & Delta == 1])
  CIF1_true2[i] <- mean(true_data1[,(Time <= ti) & Delta == 2])
  i <- i + 1 
}
```

```{r, warning = FALSE}
ggplot(tibble(time = tis, true0_vals1 = CIF0_true1, true0_vals2 = CIF0_true2,
                    true1_vals1 = CIF1_true1, true1_vals2 = CIF1_true2))+
  geom_line(aes(x = time, y = true0_vals1, color = "P1 True CIF, A0 = 0"), linewidth = 1.1, linetype = 2)+
  geom_line(aes(x = time, y = true0_vals2, color = "P2 True CIF, A0 = 0"), linewidth = 1.1, linetype = 2)+
  geom_line(aes(x = time, y = true1_vals1, color = "P1 True CIF, A0 = 1"), linewidth = 1.1, linetype = 3)+
  geom_line(aes(x = time, y = true1_vals2, color = "P2 True CIF, A0 = 1"), linewidth = 1.1, linetype = 3)+
  scale_color_manual(values = c("P1 True CIF, A0 = 0" = "pink", 
                                "P2 True CIF, A0 = 0" = "lightgreen",
                                "P1 True CIF, A0 = 1" = "hotpink",
                                "P2 True CIF, A0 = 1" = "darkgreen"))+
  labs(title = "Absolute Risk")+
  theme_bw()+
  xlim(c(0,24))+
  ylim(c(0,0.6))

ggsave("TruePopAbsRisk01.pdf")
```

We use the 100 simulated data sets of normal size to estimate the absolute risk under treatment and without treatment.

```{r}
timesRF <- list()
CIF0_vals1 <- list()
CIF0_vals2 <- list()
CIF1_vals1 <- list()
CIF1_vals2 <- list()
timesCox <- list();
AR01 <- list()
AR02 <- list()
AR11 <- list()
AR12 <- list()

for(i in 1:100){
  # RF
  timesRF[[i]] <- RF_fits[[i]]$time.interest
  CIF0_vals1[[i]] <- RF_fits[[i]]$cif[,,1][new_data[[i]]$ch == "N",] |> colMeans()
  CIF0_vals2[[i]] <- RF_fits[[i]]$cif[,,2][new_data[[i]]$ch == "N",]  |> colMeans()
  CIF1_vals1[[i]] <- RF_fits[[i]]$cif[,,1][new_data[[i]]$ch == "Y",] |> colMeans()
  CIF1_vals2[[i]] <- RF_fits[[i]]$cif[,,2][new_data[[i]]$ch == "Y",] |> colMeans()
  
  
  # Cox Without chemo
  # Linear predictors
  exp_terms1 <- exp(predict(Cox_fits1[[i]], type="lp", reference = "zero"))[new_data[[i]]$ch == "N"]
  exp_terms2 <- exp(predict(Cox_fits2[[i]], type="lp", reference = "zero"))[new_data[[i]]$ch == "N"]
  # Times
  timesCox[[i]] <- basehaz(Cox_fits1[[i]])$time
  # Cause specific cumulative hazards
  Haz1 <- exp_terms1 %*% t(basehaz(Cox_fits1[[i]], centered = FALSE)$hazard)
  Haz2 <- exp_terms2 %*% t(basehaz(Cox_fits2[[i]], centered = FALSE)$hazard)
  # Survival function
  S <- exp(-(Haz1 + Haz2))
  # Absolute risk calculation
  dLambda1 <- t(apply(Haz1, 1, function(x) c(x[1], diff(x))))
  dLambda2 <- t(apply(Haz2, 1, function(x) c(x[1], diff(x))))
  AR01[[i]] <- t(apply(S * dLambda1, 1, cumsum))
  AR01[[i]] <- colMeans(AR01[[i]])
  AR02[[i]] <- t(apply(S * dLambda2, 1, cumsum))
  AR02[[i]] <- colMeans(AR02[[i]])
  
  # Cox With chemo
  # Linear predictors
  exp_terms1 <- exp(predict(Cox_fits1[[i]], type="lp", reference = "zero"))[new_data[[i]]$ch == "Y"]
  exp_terms2 <- exp(predict(Cox_fits2[[i]], type="lp", reference = "zero"))[new_data[[i]]$ch == "Y"]
  # Times
  timesCox[[i]] <- basehaz(Cox_fits1[[i]])$time
  # Cause specific cumulative hazards
  Haz1 <- exp_terms1 %*% t(basehaz(Cox_fits1[[i]], centered = FALSE)$hazard)
  Haz2 <- exp_terms2 %*% t(basehaz(Cox_fits2[[i]], centered = FALSE)$hazard)
  # Survival function
  S <- exp(-(Haz1 + Haz2))
  # Absolute risk calculation
  dLambda1 <- t(apply(Haz1, 1, function(x) c(x[1], diff(x))))
  dLambda2 <- t(apply(Haz2, 1, function(x) c(x[1], diff(x))))
  AR11[[i]] <- t(apply(S * dLambda1, 1, cumsum))
  AR11[[i]] <- colMeans(AR11[[i]])
  AR12[[i]] <- t(apply(S * dLambda2, 1, cumsum))
  AR12[[i]] <- colMeans(AR12[[i]])
}
```

Now for the plots. Creating a plot data.frame for no treatment. 

```{r}
df_plot_notreat <- map_dfr(1:100, function(i) {
  tibble(
    time  = c(timesRF[[i]],
              timesRF[[i]],
              timesCox[[i]],
              timesCox[[i]]),
    value = c(CIF0_vals1[[i]],
              CIF0_vals2[[i]],
              AR01[[i]],
              AR02[[i]]),
    type  = c(rep("RF CIF", 2*length(timesRF[[i]])),
              rep("Cox CIF", 2*length(timesCox[[i]]))),
    process = c(rep("P1", length(timesRF[[i]])),
              rep("P2", length(timesRF[[i]])),
              rep("P1", length(timesCox[[i]])),
              rep("P2", length(timesCox[[i]]))),
    run = i
  )
})
```

Calculating mean in timepoints

```{r}
df_mean_cox01 <- matrix(nrow = 100, ncol = length(tis))
df_mean_cox02 <- matrix(nrow = 100, ncol = length(tis))
df_mean_RF01 <- matrix(nrow = 100, ncol = length(tis))
df_mean_RF02 <- matrix(nrow = 100, ncol = length(tis))

for(i in 1:100){
  df_mean_cox02[i,] <- approx(x = timesCox[[i]], y = AR02[[i]], xout = tis)$y
  df_mean_cox01[i,] <- approx(x = timesCox[[i]], y = AR01[[i]], xout = tis)$y
  df_mean_RF01[i,] <- approx(x = timesRF[[i]], y = CIF0_vals1[[i]], xout = tis)$y
  df_mean_RF02[i,] <- approx(x = timesRF[[i]], y = CIF0_vals2[[i]], xout = tis)$y
}

df_mean_cox01 <- colMeans(df_mean_cox01)
df_mean_cox02 <- colMeans(df_mean_cox02)
df_mean_RF01 <- colMeans(df_mean_RF01)
df_mean_RF02 <-colMeans(df_mean_RF02)

df_plot0_notreat <- tibble(
  time  = rep(tis, 6),                                             
  value = c(CIF0_true1, CIF0_true2, df_mean_cox01, df_mean_cox02, df_mean_RF01, df_mean_RF02),  
  type  = c(rep("True CIF", 2*length(tis)),                                           
            rep("Mean Cox CIF", 2*length(tis)),
            rep("Mean RF CIF", 2*length(tis))),
  process = rep(c(rep("P1", length(tis)), rep("P2", length(tis))), 3),
  run = 0)

df_plot_notreat <- rbind(df_plot_notreat, df_plot0_notreat)
df_plot1 <- df_plot_notreat[df_plot_notreat$process == "P1", ]
df_plot2 <- df_plot_notreat[df_plot_notreat$process == "P2", ]
```

```{r}
ggplot() +

  ## 1) simulation runs
  geom_line(
    data = df_plot1 %>% filter(type %in% c("RF CIF", "Cox CIF")),
    aes(time, value,
        color = type,
        linetype = type,
        group = interaction(type, run)),
    alpha = 0.4,
    linewidth = 0.9
  ) +

  ## 2) true + mean curves
  geom_line(
    data = df_plot1 %>% filter(type %in% c("True CIF", "Mean Cox CIF", "Mean RF CIF")),
    aes(time, value,
        color = type,
        linetype = type),
    linewidth = 1.2
  ) +

  scale_color_manual(values = c(
    "True CIF"      = "darkgreen",
    "RF CIF"        = "pink",
    "Cox CIF"       = "lightblue",
    "Mean Cox CIF"  = "blue4",
    "Mean RF CIF"   = "hotpink"
  )) +

  scale_linetype_manual(values = c(
    "True CIF"      = 1,
    "RF CIF"        = 2,
    "Cox CIF"       = 2,
    "Mean Cox CIF"  = 1,
    "Mean RF CIF"   = 1
  )) +

  labs(title = "Absolute Risk (P1 - No Chemo)", y = "Abs Risk") +
  theme_bw() +
  xlim(0, 24) +
  ylim(0, 0.6)

ggsave("estimation3.pdf", width = 9, height = 6)
```



```{r, warning = FALSE}
ggplot() +

  ## 1) simulation runs
  geom_line(
    data = df_plot2 %>% filter(type %in% c("RF CIF", "Cox CIF")),
    aes(time, value,
        color = type,
        linetype = type,
        group = interaction(type, run)),
    alpha = 0.4,
    linewidth = 0.9
  ) +

  ## 2) true + mean curves
  geom_line(
    data = df_plot2 %>% filter(type %in% c("True CIF", "Mean Cox CIF", "Mean RF CIF")),
    aes(time, value,
        color = type,
        linetype = type),
    linewidth = 1.2
  ) +

  scale_color_manual(values = c(
    "True CIF"      = "darkgreen",
    "RF CIF"        = "pink",
    "Cox CIF"       = "lightblue",
    "Mean Cox CIF"  = "blue4",
    "Mean RF CIF"   = "hotpink"
  )) +

  scale_linetype_manual(values = c(
    "True CIF"      = 1,
    "RF CIF"        = 2,
    "Cox CIF"       = 2,
    "Mean Cox CIF"  = 1,
    "Mean RF CIF"   = 1
  )) +

  labs(title = "Absolute Risk (P2- No Chemo)", y = "Abs Risk") +
  theme_bw() +
  xlim(0, 24) +
  ylim(0, 0.3)

ggsave("estimation4.pdf", width = 9, height = 6)
```
Creating a plot data.frame for treatment. 

```{r}
df_plot_treat <- map_dfr(1:100, function(i) {
  tibble(
    time  = c(timesRF[[i]],
              timesRF[[i]],
              timesCox[[i]],
              timesCox[[i]]),
    value = c(CIF1_vals1[[i]],
              CIF1_vals2[[i]],
              AR11[[i]],
              AR12[[i]]),
    type  = c(rep("RF CIF", 2*length(timesRF[[i]])),
              rep("Cox CIF", 2*length(timesCox[[i]]))),
    process = c(rep("P1", length(timesRF[[i]])),
              rep("P2", length(timesRF[[i]])),
              rep("P1", length(timesCox[[i]])),
              rep("P2", length(timesCox[[i]]))),
    run = i
  )
})
```

Calculating mean in timepoints

```{r}
df_mean_cox11 <- matrix(nrow = 100, ncol = length(tis))
df_mean_cox12 <- matrix(nrow = 100, ncol = length(tis))
df_mean_RF11 <- matrix(nrow = 100, ncol = length(tis))
df_mean_RF12 <- matrix(nrow = 100, ncol = length(tis))

for(i in 1:100){
  df_mean_cox12[i,] <- approx(x = timesCox[[i]], y = AR12[[i]], xout = tis)$y
  df_mean_cox11[i,] <- approx(x = timesCox[[i]], y = AR11[[i]], xout = tis)$y
  df_mean_RF11[i,] <- approx(x = timesRF[[i]], y = CIF1_vals1[[i]], xout = tis)$y
  df_mean_RF12[i,] <- approx(x = timesRF[[i]], y = CIF1_vals2[[i]], xout = tis)$y
}

df_mean_cox11 <- colMeans(df_mean_cox11)
df_mean_cox12 <- colMeans(df_mean_cox12)
df_mean_RF11 <- colMeans(df_mean_RF11)
df_mean_RF12 <-colMeans(df_mean_RF12)

df_plot0_treat <- tibble(
  time  = rep(tis, 6),                                             
  value = c(CIF1_true1, CIF1_true2, df_mean_cox11, df_mean_cox12, df_mean_RF11, df_mean_RF12),  
  type  = c(rep("True CIF", 2*length(tis)),                                           
            rep("Mean Cox CIF", 2*length(tis)),
            rep("Mean RF CIF", 2*length(tis))),
  process = rep(c(rep("P1", length(tis)), rep("P2", length(tis))), 3),
  run = 0)

df_plot_treat <- rbind(df_plot_treat, df_plot0_treat)
df_plot1 <- df_plot_treat[df_plot_treat$process == "P1", ]
df_plot2 <- df_plot_treat[df_plot_treat$process == "P2", ]
```

```{r}
ggplot() +

  ## 1) simulation runs
  geom_line(
    data = df_plot1 %>% filter(type %in% c("RF CIF", "Cox CIF")),
    aes(time, value,
        color = type,
        linetype = type,
        group = interaction(type, run)),
    alpha = 0.4,
    linewidth = 0.9
  ) +

  ## 2) true + mean curves
  geom_line(
    data = df_plot1 %>% filter(type %in% c("True CIF", "Mean Cox CIF", "Mean RF CIF")),
    aes(time, value,
        color = type,
        linetype = type),
    linewidth = 1.2
  ) +

  scale_color_manual(values = c(
    "True CIF"      = "darkgreen",
    "RF CIF"        = "pink",
    "Cox CIF"       = "lightblue",
    "Mean Cox CIF"  = "blue4",
    "Mean RF CIF"   = "hotpink"
  )) +

  scale_linetype_manual(values = c(
    "True CIF"      = 1,
    "RF CIF"        = 2,
    "Cox CIF"       = 2,
    "Mean Cox CIF"  = 1,
    "Mean RF CIF"   = 1
  )) +

  labs(title = "Absolute Risk (P1 - Chemo)", y = "Abs Risk") +
  theme_bw() +
  xlim(0, 24) +
  ylim(0, 0.6)

ggsave("estimation5.pdf", width = 9, height = 6)
```

```{r, warning = FALSE}
ggplot() +

  ## 1) simulation runs
  geom_line(
    data = df_plot2 %>% filter(type %in% c("RF CIF", "Cox CIF")),
    aes(time, value,
        color = type,
        linetype = type,
        group = interaction(type, run)),
    alpha = 0.4,
    linewidth = 0.9
  ) +

  ## 2) true + mean curves
  geom_line(
    data = df_plot2 %>% filter(type %in% c("True CIF", "Mean Cox CIF", "Mean RF CIF")),
    aes(time, value,
        color = type,
        linetype = type),
    linewidth = 1.2
  ) +

  scale_color_manual(values = c(
    "True CIF"      = "darkgreen",
    "RF CIF"        = "pink",
    "Cox CIF"       = "lightblue",
    "Mean Cox CIF"  = "blue4",
    "Mean RF CIF"   = "hotpink"
  )) +

  scale_linetype_manual(values = c(
    "True CIF"      = 1,
    "RF CIF"        = 2,
    "Cox CIF"       = 2,
    "Mean Cox CIF"  = 1,
    "Mean RF CIF"   = 1
  )) +

  labs(title = "Absolute Risk (P2 - Chemo)", y = "Abs Risk") +
  theme_bw() +
  xlim(0, 24) +
  ylim(0, 0.3)

ggsave("estimation6.pdf", width = 9, height = 6)
```

