---
title: 'Hidden Confounder'
output: html_document
date: "2024-10-18"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, include=FALSE}
# Libraries
library(data.table)
library(survival)
#install.packages("rtmle")
#devtools::install_github("tagteam/rtmle")
library(rtmle)
library(ggplot2)
library(microbenchmark)
theme_set(theme_bw())
source("~/Helene_projekt/funcs_for_two_settings.R")
```

# Simulation function

```{r}
set.seed(2964)
set1_data <- sim_data_setting1(N = 5000)
set1_data_fu <- sim_data_setting1(N = 10000, followup = 5)
plot_event_data(set1_data)
plot_event_data(set1_data_fu)
```

# Proportions

```{r}
summary_sim_event(set1_data)
summary_sim_event(set1_data_fu)
```

# Fitting model

```{r}
set1_data_t <- trans_int_data(set1_data)

set1_data_t[, at_risk_oper := as.numeric(A == 0)]
set1_data_t[, at_risk_cov := as.numeric(L == 0)]

survfit_oper <- coxph(Surv(tstart, tstop, Delta == 0) ~ L0 + L, 
                      data = set1_data_t[at_risk_oper == 1], cluster = ID)
survfit_death <- coxph(Surv(tstart, tstop, Delta == 1) ~ L0 + A + L, 
                      data = set1_data_t, cluster = ID)
survfit_cens <- coxph(Surv(tstart, tstop, Delta == 2) ~ L0 + A + L, 
                      data = set1_data_t, cluster = ID)
survfit_cov <- coxph(Surv(tstart, tstop, Delta == 3) ~ L0 + A, 
                    data = set1_data_t[at_risk_cov == 1], cluster = ID)

survfit_oper$coefficients 
survfit_death$coefficients 
survfit_cens$coefficients 
survfit_cov$coefficients
```

Den estimere ikke rigtigt når vi har censorering...

```{r}
set1_data_t <- trans_int_data(set1_data_fu)
set1_data_t[, C := !is.na(Delta)]

set1_data_t[, at_risk_oper := as.numeric(A == 0)]
set1_data_t[, at_risk_cov := as.numeric(L == 0)]

survfit_oper <- coxph(Surv(tstart, tstop, Delta == 0) ~ L0 + L, 
                      data = set1_data_t[at_risk_oper == 1], cluster = ID)
survfit_death <- coxph(Surv(tstart, tstop, Delta == 1) ~ L0 + A + L, 
                      data = set1_data_t, cluster = ID)
survfit_cens <- coxph(Surv(tstart, tstop, Delta == 2) ~ L0 + A + L, 
                      data = set1_data_t, cluster = ID)
survfit_cov <- coxph(Surv(tstart, tstop, Delta == 3) ~ L0 + A, 
                    data = set1_data_t[at_risk_cov == 1], cluster = ID)

survfit_oper$coefficients 
survfit_death$coefficients 
survfit_cens$coefficients 
survfit_cov$coefficients
```

Note that we can affect the intensities by:

$$\lambda^x_k(t\, \vert \, \mathcal{F}_{t-}) =  R^x(t,k)  \eta^x \nu^x t^{\nu^x - 1} 
 \exp(\beta^T Z)$$

changing the value of $\eta^x$. For a follow up time $K$ 

```{r}
eta_vals <- seq(0.05, 0.15, by = 0.01)
etas <- expand.grid(eta1 = eta_vals, eta2 = 0.08, eta3 = 0.08, eta4 = 0.08)

K <- 5
summaries <- matrix(ncol = 7, nrow = nrow(etas))

for(i in 1:nrow(etas)){
  print(i)
  eta <- etas[i,]
  sim <- sim_data_setting1(1000, eta = eta, followup = K)
  summaries[i,] <- summary_sim_event(sim)
}

colnames(summaries) <- names(summary_sim_event(sim))
head(summaries)

p1 <- ggplot(data = summaries)+
  geom_line(aes(x = etas[,1], y = `Prop Op`, color = "Op"))+
  geom_line(aes(x = etas[,1], y = `Prop Cov`, color = "Cov"))+
  geom_line(aes(x = etas[,1], y = `Prop Death`, color = "Death"), linetype = 2)+
  geom_line(aes(x = etas[,1], y = `Prop Cens`, color = "Cens"), linetype = 2)+
  geom_line(aes(x = etas[,1], y = `Prop FU Cens`, color = "FU Cens"), linetype = 2)+
  scale_color_manual(values = c("red3", "blue2", "green3", "purple3", "pink"))

p1


#p2 <- ggplot(data = summaries)+
#  geom_line(aes(x = etas[,4], y = `Prop Op`, color = "Op"))+
#  geom_line(aes(x = etas[,4], y = `Prop Cov`, color = "Cov"))+
#  geom_line(aes(x = etas[,4], y = `Prop Death`, color = "Death"), linetype = 2)+
#  geom_line(aes(x = etas[,4], y = `Prop Cens`, color = "Cens"), linetype = 2)+
#  geom_line(aes(x = etas[,4], y = `Prop FU Cens`, color = "FU Cens"), linetype = 2)+
#  scale_color_manual(values = c("red3", "blue2", "green3", "purple3", "pink"))
#
#gridExtra::grid.arrange(p1, p2, ncol = 2)
```

# Investigating model misspecifications

Function for fitting and extracting model estimates, when the model is specified correctly, misspecified in terms of missing time varying covariate effect and misspecified in terms of missing operation effect.  

```{r}
fit_mod_set1(set1_data)
```
Analysing error in modelsspecifications with repeated simulations. 

```{r}
# TODO: inkorporer followup i den her
set.seed(9457)
try1 <- sys_invest(N = 400, eff_A_D = -1, timefix = TRUE)
try2 <- sys_invest(N = 400, eff_L_A = 2, eff_L_D = 2, eff_A_D = -1, timefix = FALSE)
try3 <- sys_invest(N = 400, eff_L_A = 3, eff_L_D = 3, eff_A_D = -1, timefix = FALSE) # This runs only with the timefix = FALSE
```

Histograms of the estimates when the models are misspecified:

```{r}
try1$pp3
try2$pp3
try3$pp3 # Jeg tror at modellerne nogle gange fejler og estimerer Cens L meget stort...?
# Det er mærkeligt at plotet ikke kommer frem, estimaterne er der, hvis man fjerner de to meget ekstreme estimater
# er bias heller ikke så stort

#jpeg(filename = "hidden_confounder.jpg", width = 600)
#try3$pp3
#dev.off()
```

Bias:

```{r}
try1$Bias
try2$Bias
try3$Bias
```

Der er bias i dødsprocessen (hidden confounder).

Confidence intervals:

```{r}
quant_data <- try1$data
CIs <- quant_data[, .(Q2.5 = quantile(`Fully Specified`, 0.025),
               Q97.5 = quantile(`Fully Specified`, 0.975)), by = "Est"]

# The estimates are very spread out...
CIs
# But the 95 confidence intervals contain the true parameters
(CIs[,2] <= quant_data[1:nrow(CIs),True])  & (quant_data[1:nrow(CIs),True]  <= CIs[,3])

# With the hidden confounder, this is not the case (at least in the extreme case)
quant_data <- try3$data
CIs <- quant_data[, .(Q2.5 = quantile(`-L`, 0.025, na.rm = TRUE),
               Q97.5 = quantile(`Fully Specified`, 0.975, na.rm = TRUE)), by = "Est"]

(CIs[,2] <= quant_data[1:nrow(CIs),True])  & (quant_data[1:nrow(CIs),True]  <= CIs[,3])
```

## Why do we get the errors???

Example of errors

```{r}
data2 <- sim_data_setting1(N = 20000, beta_L_A = 2, beta_L_D = 3, beta_A_D = -0.5,
                           nu = rep(1.3,4), eta= rep(0.01,4))
data1 <- sim_data_setting1(N = 20000, beta_L_A = 2, beta_L_D = 3, beta_A_D = -0.5)

# when we fit a model here, we get errors
trans_data <- trans_int_data(data1)
trans_data[, at_risk_oper := as.numeric(A == 0)]
trans_data[, at_risk_cov := as.numeric(L == 0)]

survfit_oper <- coxph(Surv(tstart, tstop, Delta == 0) ~ L0 + L,
                      data = trans_data[at_risk_oper == 1], cluster = ID)
survfit_death <- coxph(Surv(tstart, tstop, Delta == 1) ~ L0 + A + L,
                       data = trans_data, cluster = ID)
survfit_cens <- coxph(Surv(tstart, tstop, Delta == 2) ~ L0 + A + L,
                      data = trans_data, cluster = ID)
survfit_cov <- coxph(Surv(tstart, tstop, Delta == 3) ~ L0 + A,
                     data = trans_data[at_risk_cov == 1], cluster = ID)
coefficients(survfit_oper); coefficients(survfit_death); coefficients(survfit_cens); coefficients(survfit_cov) 
```

We examine data
```{r}
summary_sim_event(data1)
```
Looks quite reasonable. We look at the length of the intervals
```{r}
newdata <- trans_data[, .(intlength= (tstop - tstart))]

# Der er nogle intervaller som er meget små
shorties <- which(newdata$intlength <= 10^(-3))
newdata[shorties,] # de er helt vildt små! (e-15)
trans_data[shorties, Delta] |> as.factor() |> summary() # Flest Type 1 events (død)
trans_data[shorties, table(Delta,at_risk_cov)]
trans_data[shorties, k] |> as.factor() |> summary() # Og flest når det er event nummer 2
trans_data[shorties, at_risk_oper] |> as.factor() |> summary() # og når der ikke har været en operation
trans_data[shorties, at_risk_cov] |> as.factor() |> summary() # og når der har været et covariate skifte
```
The interval lengths are incredibly small...

```{r}
newdata$intlength |> mean() 
newdata$intlength |> quantile(c(0.025, 0.975)) 
```
We try to remove the very small intervals
```{r}
trans_data_minusobs <- trans_data[-shorties,]


trans_data_minusobs[, at_risk_oper := as.numeric(A == 0)]
trans_data_minusobs[, at_risk_cov := as.numeric(L == 0)]

# True models:
survfit_oper <- coxph(Surv(tstart, tstop, Delta == 0) ~ L0 + L,
                      data = trans_data_minusobs[at_risk_oper == 1], cluster = ID)
survfit_death <- coxph(Surv(tstart, tstop, Delta == 1) ~ L0 + A + L,
                       data = trans_data_minusobs, cluster = ID)
survfit_cens <- coxph(Surv(tstart, tstop, Delta == 2) ~ L0 + A + L,
                      data = trans_data_minusobs, cluster = ID)
survfit_cov <- coxph(Surv(tstart, tstop, Delta == 3) ~ L0 + A,
                     data = trans_data_minusobs[at_risk_cov == 1], cluster = ID)

survfit_oper$coefficients; survfit_death$coefficients; survfit_cens$coefficients; survfit_cov$coefficients
```

Hvis vi fjerner de meget meget små intervaller, så går det godt... Med timefix = FALSE får vi estimaterne:

```{r}
survfit_oper1 <- coxph(Surv(tstart, tstop, Delta == 0) ~ L0 + L, timefix = FALSE,
                      data = trans_data[at_risk_oper == 1], cluster = ID)
survfit_death1 <- coxph(Surv(tstart, tstop, Delta == 1) ~ L0 + A + L,
                       data = trans_data, cluster = ID, timefix = FALSE)
survfit_cens1 <- coxph(Surv(tstart, tstop, Delta == 2) ~ L0 + A + L,
                      data = trans_data, cluster = ID, timefix = FALSE)
survfit_cov1 <- coxph(Surv(tstart, tstop, Delta == 3) ~ L0 + A, timefix = FALSE,
                     data = trans_data[at_risk_cov == 1], cluster = ID)

survfit_oper1$coefficients; survfit_death1$coefficients; survfit_cens1$coefficients; survfit_cov1$coefficients
```
Man får lidt forskellige, men meget ens resultater, men vi har også virkelig mange observationer...

Plot of intensities, i base case ish:
```{r}
beta <- matrix(ncol = 4, nrow = 4)
# A0 is 0
beta[3,] <- 0
# L0 increases the probability of L = 1 and A = 1
beta[1,c(1,4)] <- c(1,1)
# L=1 increases the probability of A = 1
beta[4,1] <- 2
# A=1 decreases the risk of death
beta[2,2] <- -1
# L0,L=1 increases the risk of death
beta[c(1,4),2] <- c(1, 2)
# Censorering does not depend on anything
beta[,3] <- 0

# A = 1 does not affect the intensity of A (the event occurs only once)
beta[2,1] <- 0
# L = 1 does not affect the intensity of L (the event occurs only once)
beta[4,4] <- 0
# My assumption: A = 1 decreases the risk of L = 1
beta[2,4] <- -0.5

nu <- 1.1
eta <- 0.1

A <- 1
L <- 1
L0 <- 0.5

lambda_A <- function(t) {
  eta * nu * t ^ (nu - 1) * exp(L0 * beta[1,1] + A * beta[2,1] + L* beta[4,1])
}
lambda_D <- function(t) {
  eta * nu * t ^ (nu - 1) * exp(L0 * beta[1,2] + A * beta[2,2] + L* beta[4,2])
}
lambda_C <- function(t) {
  eta * nu * t ^ (nu - 1) * exp(L0 * beta[1,3] + A * beta[2,3] + L* beta[4,3])
}
lambda_L <- function(t) {
  eta * nu * t ^ (nu - 1) * exp(L0 * beta[1,4] + A * beta[2,4] + L* beta[4,4])
}

ggplot()+
  geom_function(fun = lambda_A, aes(color = "A"))+
  geom_function(fun = lambda_D, aes(color = "D"))+
  geom_function(fun = lambda_C, aes(color = "C"))+
  geom_function(fun = lambda_L, aes(color = "L"))+
  scale_color_manual(values = c("A" = "red4", "D" = "blue3", "C" = "lightblue3", "L" = "purple2"))
```

```{r}
L <- 0
lambda_nu_high <- function(t) {
  eta * 1.5 * t ^ (1.5 - 1) * exp(L0 * beta[1,1] + A * beta[2,1] + L* beta[4,1])
}
lambda_nu_low <- function(t) {
  eta * 1.05 * t ^ (1.05 - 1) * exp(L0 * beta[1,1] + A * beta[2,1] + L* beta[4,1])
}
lambda_eta_high <- function(t) {
  0.3 * nu * t ^ (nu - 1) * exp(L0 * beta[1,3] + A * beta[2,3] + L* beta[4,3])
}
lambda_eta_low <- function(t) {
  0.01 * nu * t ^ (nu - 1) * exp(L0 * beta[1,4] + A * beta[2,4] + L* beta[4,4])
}

ggplot()+
  geom_function(fun = lambda_nu_high, aes(color = "nu high"))+
  geom_function(fun = lambda_nu_low, aes(color = "nu low"))+
  geom_function(fun = lambda_eta_high, aes(color = "eta high"))+
  geom_function(fun = lambda_eta_low, aes(color = "eta low"))+
  scale_color_manual(values = c("nu high" = "red4", "nu low" = "blue3", "eta high" = "lightblue3", "eta low" = "purple2"))
```

Umiddelbart ser det mest fornuftigt ud hvis vi vælger $\eta = 0.01$ og $\nu = 1.5$.


```{r}
lambda_reasonable <- function(t) {
  0.01 * 1.4 * t ^ (1.3 - 1) * exp(L0 * beta[1,4] + A * beta[2,4] + L* beta[4,4])
}

ggplot()+
  geom_function(fun = lambda_reasonable, aes(color = "Reasonable"))+
  scale_color_manual(values = c("Reasonable" = "red4"))
```




