---
title: "Effect of interventions"
output: html_document
date: "2024-11-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Libraries
library(gridExtra)
library(tidyverse)
library(survival)
source("~/Helene_projekt/Source code/sim_surv.R")
source("~/Helene_projekt/Source code/Simple_functions.R")
```


# The simple function

## 1. Simulate data

```{r}
n = 10^4
sim1 <- simple1(2000, c(1,2,3), c(0.1,0.1,0.1), c(1.1,1.1,1.1))
```


## 2. Fit a Weibull 

```{r}
survfit <- survreg(Surv(Time, Delta == 1) ~ 1, data = sim1, dist='weibull')
```

## 3. Estimate parameters

```{r}
shape_est <- 1/survfit$scale; 
shape_est; 1.1
scale_est <- 1/exp(survfit$coefficients[1]);
scale_est; 0.1
```

## 4. Simulate new data using the estimated parameters

```{r}
sim_obs <- simple1(n, c(0,0,0), rep(scale_est, 3), rep(shape_est, 3))
```

## 5. Simulate new data under intervention

We intervene by setting $L0 = 0$

```{r}
sim_int <- simple1(n, rep(0,3), rep(scale_est, 3), rep(shape_est, 3))
```

## 6. Estimate average causal effects

```{r}
mean(sim_obs[sim_obs$Delta == 0,]$Time) - mean(sim_int[sim_int$Delta == 0,]$Time)
mean(sim_obs[sim_obs$Delta == 1,]$Time) - mean(sim_int[sim_int$Delta == 1,]$Time)
mean(sim_obs[sim_obs$Delta == 2,]$Time) - mean(sim_int[sim_int$Delta == 2,]$Time)
```
# An intervention on treatment

## 1. Simulate data

```{r}
n = 10^4
eta <- c(0.1, 0.1, 0.1, 0.1)
nu <- c(1.1, 1.1, 1.1, 1.1)
# Effect on Operation
beta0 <- c(3, 0, 1, 9)
# Effect on Censoring
beta1 <- c(0, 0, -1, 5)
# Effect on Death
beta2 <- c(3, 0, 1, 1)
# Effect on time varying covariate
beta3 <- c(3, 0, -1, 0.5)
beta <- cbind(beta0, beta1, beta2, beta3)
term_deltas <- c(1, 2)

at_risk2 <- function(x, k, m) {
  # If you have not died yet or been censored yet, you are at risk for dying or being censored
  if(x == 1 | x == 2) return(1)
  # You are only at risk for an operation if you have not had an operation yet
  else if(x == 0) return(as.numeric(k == 0 | (k == 1 & m == 1)))
  # You are only at risk for a change in the covariate process if you have not had a change yet
  else return(as.numeric(m == 0))
}

sim1 <- sur_sim(4000, beta, eta, nu, at_risk2, term_deltas) %>%
  mutate(k = ave(ID, ID, FUN = seq_along) - 1)
```


## 2. Fit a Weibull 

```{r}
survfit0 <- survreg(Surv(Time, Delta == 0) ~ L0 + A + L1, data = sim1[sim1$k == 0,], dist='weibull')
survfit1 <- survreg(Surv(Time, Delta == 1) ~ L0 + A + L1, data = sim1[sim1$k == 0,], dist='weibull')
survfit2 <- survreg(Surv(Time, Delta == 2) ~ L0 + A + L1, data = sim1[sim1$k == 0,], dist='weibull')
survfit3 <- survreg(Surv(Time, Delta == 3) ~ L0 + A + L1, data = sim1[sim1$k == 0,], dist='weibull')
```

## 3. Estimate parameters

```{r}
nu_est <- c(1/survfit0$scale, 1/survfit1$scale, 1/survfit2$scale, 1/survfit3$scale) 
nu_est; 1.1
eta_est <- c(1/exp(survfit0$coefficients[1]), 1/exp(survfit1$coefficients[1]), 
             1/exp(survfit2$coefficients[1]), 1/exp(survfit3$coefficients[1]))
eta_est; 0.1
```

## 4. Simulate new data using the estimated parameters

```{r}
sim_obs <- sur_sim(N = n, beta = beta, eta = eta_est, nu = nu_est, at_risk = at_risk2,
                   term_deltas = term_deltas)
```

## 5. Simulate new data under intervention

We intervene by setting the effect of treatment to 0. 

```{r}
betas_int <- beta
betas_int[3,] <- rep(0,4)

sim_int <- sur_sim(N = n, beta = betas_int, eta = eta_est, nu = nu_est, at_risk = at_risk2,
                   term_deltas = term_deltas)
```

## 6. Estimate average causal effects

```{r}
mean(sim_obs[sim_obs$Delta == 0 & sim_obs$A == 1,]$Time) - mean(sim_int[sim_int$Delta == 0,]$Time)
mean(sim_obs[sim_obs$Delta == 1 & sim_obs$A == 1,]$Time) - mean(sim_int[sim_int$Delta == 1,]$Time)
mean(sim_obs[sim_obs$Delta == 2 & sim_obs$A == 1,]$Time) - mean(sim_int[sim_int$Delta == 2,]$Time)
```
