---
title: 'Simulation study: Cause Specific Hazards'
output: html_document
date: "2024-10-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Libraries
library(gridExtra)
library(tidyverse)
```


# Competing risk

Simulation of covariates to time 0:

```{r}
n <- 1700

X0 <- rbinom(n, 1, 0.1627)
X1 <- rpois(n, 35)
L0 <- cbind(X0, X1)

# Effekt
alpha <- c(0, 0)
beta <- c(0.5816, -0.1)
```

Hazard functions

```{r}
# Parameters
lambda <- 0.0271
v <- 1.3841

# Specify hazard functions
weib_haz <- function(t) lambda * v * t ^ (v - 1)
weib_cum_haz <- function(t) lambda * t ^ v 

# Cause specific hazard functions
haz1 <- function(t) weib_haz(t) * exp(L0 %*% alpha)
haz2 <- function(t) weib_haz(t) * exp(L0 %*% beta)
```

Distribution function and inverse distribution function

```{r}
# CDF of time to any event T
F_func <- function(t, x) {
  1 - exp( - (weib_cum_haz(t) * exp(t(alpha) %*% x) + 
            weib_cum_haz(t) * exp(t(beta) %*% x)))
}

# Inverse of the distribution function

inverse_F <- function(p, x, lower_bound = 0, upper_bound = 100) {
  # Define the root-finding function (F(t) - p)
  root_function <- function(t) {
    F_func(t, x) - p
  }
  # Use uniroot to find the value of t such that F(t) = y
  result <- uniroot(root_function, lower = lower_bound, upper = upper_bound)
  
  # Return the value of t
  return(result$root)
}
```

Simulating time to event variables

```{r}
# Simulating uniform random variables
Z <- runif(n)
# Simulating event times
Ts <-  numeric(n)
for(i in 1:n){
  Ts[i] <- inverse_F(p = Z[i], x = L0[i,])
}

# Simulating censoring
censoring_time <- rweibull(n, 1.3841, 1/0.0271)
# The event variable
event <- ifelse(Ts <= censoring_time, 1, 0)
```

Simulating the cause

```{r}
# The probability of cause of event being 1
p1 <- haz1(Ts) / (haz1(Ts) + haz2(Ts))

cause_type <- ifelse(event == 1, rbinom(n, 1, 1 - p1) + 1, 0) # 1 indicates cause 1, 2 indicates cause of event being 2

# Cause specific event times
T1 <- Ts[cause_type == 1]
T2 <- Ts[cause_type == 2]
```

Plots of distribution of evnt times

```{r}
plot_1 <- ggplot(data = data.frame(T1 = T1), aes(x = T1, y = ..density..)) + 
  geom_histogram(binwidth = 1, fill = "blue", alpha = 0.5) + 
  theme_minimal()+
  labs(title = "Cause 1")

plot_2 <- ggplot(data = data.frame(T2 = T2), aes(x = T2, y = ..density..)) + 
  geom_histogram(binwidth = 1, fill = "blue", alpha = 0.5) + 
  theme_minimal()+
  labs(title = "Cause 2")

grid.arrange(plot_1, plot_2, ncol = 2)
```

Output of function

```{r}
tibble(
  Id = n,
  X0 = X0,
  X1 = X1,
  "Time to event" = Ts,
  "Cause" = cause_type
)
```


# Simulating data with a non-terminal event

We simulate with number of individuals

```{r}
N <- 100
```

Shape and scale parameter

```{r}
eta <- 0.1
nu <- 1.2
```

Intensities (for $x = a,d,c$)

```{r}
# Effects
alpha <- c("a" = 0.1, "d" = 0.2, "c" = 0.3)


lambda <- function(t, x, k) {
  eta * nu * t ^ (nu - 1) * phi(t, x, k)
}

phi <- function(t, x, k) {
  if(x == "a" & k == 2){
    return(0)
  }
  exp(L0 * alpha[x])
}
```

Summed cumulative hazard for waiting times

```{r}
sum_cum_haz <- function(u, t, k) {
  term <- eta * (phi(t,"a",k) + phi(t, "d", k) + phi(t, "c", k))
  term * ((t + u) ^ nu - t ^ nu)
}

sum_cum_haz_inv <- function(u, t, k) {
  term <- eta * (phi(t,"a",k) + phi(t, "d", k) + phi(t, "c", k))
  ((u + term * t ^ nu) / term ) ^ (1/nu) - t
}
```

Probabilities of event type

```{r}
probs <- function(t, k){
  summ <- phi(t,"a",k) + phi(t, "d", k) + phi(t, "c", k)
  c(phi(t, "a", k) / summ, phi(t, "d", k) / summ, phi(t, "c", k) / summ)
} 
```

Simulation

```{r}
res <- tibble(ID = numeric(), L0 = numeric(), Time = numeric(), Delta = numeric())

set.seed(98)

for(j in 1:N){
  
  # Draw
  L0 <- runif(1)
  
  # Initialize
  T_k1 <- 0
  Delta <- 0
  
  # Iterate
  
  Ts <- c()
  Deltas <- c()
  k <- 1
  
  while(Delta == 0){
    V <- runif(1)
    W_k1 <- sum_cum_haz_inv(-log(V), L0, k)
    T_k1 <- T_k1 + W_k1
    Ts[k] <- T_k1
    Deltas[k] <- sample(c(0,1,2), size = 1, prob = probs(T_k1, k))
    print(probs(T_k1, k))
    Delta <- Deltas[k]
    k <- k + 1
  }
  
  jth_res <- tibble(ID = rep(j, length(Ts)), 
                    L0 = rep(L0,length(Ts)), 
                    Time = Ts, 
                    Delta = Deltas) 
  res <- bind_rows(res, jth_res)
}

res

table(res$Delta)
```

# Writing a simulation function

We create a simulation function that is a function of number of individuals $N$, the effect of the baseline covariate on the different intensities $\alpha \in \mathbb{R}^3$, the shape and scale parameter $\eta$ and $\nu$. 

```{r}
N <- 100
eta <- 0.1
nu <- 1.2
alpha <- c("a" = 0.1, "d" = 0.2, "c" = 0.3)

sur_sim <- function(N, alpha, eta, nu) {
  
  # Intensities

  phi <- function(t, x, k) {
    if(x == "a" & k == 2){
      return(0)
    }
    exp(L0 * alpha[x])
  }
  
  # Inverse Summed Cumulative hazard

  sum_cum_haz_inv <- function(u, t, k) {
    term <- eta * (phi(t,"a",k) + phi(t, "d", k) + phi(t, "c", k))
    ((u + term * t ^ nu) / term ) ^ (1/nu) - t
  }
  
  # Event probabilities
  probs <- function(t, k){
    summ <- phi(t,"a",k) + phi(t, "d", k) + phi(t, "c", k)
    c(phi(t, "a", k) / summ, phi(t, "d", k) / summ, phi(t, "c", k) / summ)
  }
  
  # Simulation
  res <- tibble(ID = numeric(), L0 = numeric(), Time = numeric(), Delta = numeric())

  for(j in 1:N){
    
    # Draw
    L0 <- runif(1)
    
    # Initialize
    T_k1 <- 0
    Delta <- 0
    
    # Iterate
    Ts <- c()
    Deltas <- c()
    k <- 1
    
    while(Delta == 0){
      V <- runif(1)
      W_k1 <- sum_cum_haz_inv(-log(V), T_k1, k)
      T_k1 <- T_k1 + W_k1
      Ts[k] <- T_k1
      Deltas[k] <- sample(c(0,1,2), size = 1, prob = probs(T_k1, k))
      Delta <- Deltas[k]
      k <- k + 1
    }
    
    jth_res <- tibble(ID = rep(j, length(Ts)), 
                      L0 = rep(L0,length(Ts)), 
                      Time = Ts, 
                      Delta = Deltas) 
    res <- bind_rows(res, jth_res)
  }
  return(res)
}
```

Trying out the new simulation function

```{r}
sur_sim(N, alpha, eta, nu)
```


