---
title: 'T2D'
output: html_document
date: "2024-10-18"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Libraries
library(data.table)
library(survival)
#install.packages("rtmle")
#devtools::install_github("tagteam/rtmle")
library(rtmle)
library(ggplot2)
library(microbenchmark)
theme_set(theme_bw())
source("~/Helene_projekt/funcs_for_two_settings.R")
```

# The Simulation Function

```{r}
set.seed(2964)
N <- 10000
set2a_data <- sim_data_setting2(N = N, beta_L_D = 1, cens = 0) # T2D has effect on Death and we have no censoring
set2b_data <- sim_data_setting2(N = N, beta_L_D = 0, cens = 0)# T2D has no effect on Death and we have no censoring
set2c_data <- sim_data_setting2(N = N, beta_L_D = 0.5, cens = 0)# T2D has small effect on Death and we have no censoring
set2a_fu_data <- sim_data_setting2(N = N, beta_L_D = 1, followup = 5)
set2b_fu_data <- sim_data_setting2(N = N, beta_L_D = 0, followup = 5)
```


```{r}
plot_event_data(set2a_data)
plot_event_data(set2a_fu_data)
plot_event_data(set2b_data)
plot_event_data(set2b_fu_data)
```


# Proportions

```{r}
summary_sim_event(set2a_data)
summary_sim_event(set2a_fu_data)
summary_sim_event(set2b_data)
summary_sim_event(set2b_fu_data)
```

# Fitting models

```{r}
# Transform data into tstart tstop format
mod_data_a <- trans_int_data(set2a_fu_data)

mod_data_a[, at_risk_cov := as.numeric(L == 0)]

# Fit models
survfit_death <- coxph(Surv(tstart, tstop, Delta == 1) ~ I(L0 / 50) + A0 + L + L1, 
                       data = mod_data_a, cluster = ID)
survfit_cens <- coxph(Surv(tstart, tstop, Delta == 0) ~ I(L0 / 50) + A0 + L + L1, 
                      data = mod_data_a, cluster = ID)
survfit_cov <- coxph(Surv(tstart, tstop, Delta == 3) ~ I(L0 / 50) + A0 + L1, 
                     data = mod_data_a[at_risk_cov == 1], cluster = ID)

survfit_death$coefficients 
survfit_cens$coefficients 
survfit_cov$coefficients
```

# Finding effect of T2D on Death

We want to find

$$\frac{P(X(\tau) = 2)}{P(X(\tau) = 0)}$$
And 
$$\frac{P(X(\tau) = 3)}{P(X(\tau) = 1)}$$
Where $\tau \in \mathbb{R}$. That is the fraction of the probability of being dead without T2D at time $\tau$ over the probability of not having T2D at time $\tau$ and the fraction of the probability of being dead with T2D at time $\tau$ over the probability of having T2D at time $\tau$. We define $\tau$ as 

```{r}
tau <- 5
```

```{r}
prob_fracs(set2a_data) # In the data where the change in the covariate process (L = 1) increases risk of death
prob_fracs(set2b_data) # In the data where the change in the covariate process does not increase risk of death
prob_fracs(set2c_data) # In the data where the change in the covariate process increases risk of death a little bit
```

Repeated calculations:

```{r}
set.seed(8453)
B <- 10 
N <- 12000
results <- matrix(NA, nrow = B, ncol = 6)
results10 <- matrix(NA, nrow = B, ncol = 6)
colnames(results) <- c("A: WO T2D", "A: T2D", "B: WO T2D", "B: T2D", "C: WO T2D", "C: T2D")
colnames(results10) <- c("A: WO T2D", "A: T2D", "B: WO T2D", "B: T2D", "C: WO T2D", "C: T2D")

for(i in 1:B){
  print(i)
  data_A <- sim_data_setting2(N = N, beta_L_D = 1, cens = 0) 
  data_B <- sim_data_setting2(N = N, beta_L_D = 0, cens = 0)
  data_C <- sim_data_setting2(N = N, beta_L_D = 0.5, cens = 0)
  results[i, 1:2] <- prob_fracs(data_A) 
  results[i, 3:4] <- prob_fracs(data_B)
  results[i, 5:6] <- prob_fracs(data_C)
  results10[i, 1:2] <- prob_fracs(data_A, tau = 10) 
  results10[i, 3:4] <- prob_fracs(data_B, tau = 10)
  results10[i, 5:6] <- prob_fracs(data_C, tau = 10)
}
```


```{r}
dataa <- data.table(apply(results, 2, FUN = roll_avg))
results <- data.table(results)
results[, i := 1:B]
dataa[, i := 1:B]

dataa10 <- data.table(apply(results10, 2, FUN = roll_avg))
results10 <- data.table(results10)
results10[, i := 1:B]
dataa10[, i := 1:B]
```

```{r}
#jpeg("T2D_5.jpg")
ggplot(melt(results, measure.vars = 1:6))+
  geom_line(aes(x = i, y = value, col = variable), size = 1, linetype = "dashed")
#dev.off()

ggplot(melt(dataa, measure.vars = 1:6))+
  geom_line(aes(x = i, y = value, col = variable), size = 1, linetype = "dashed")
```
All scenarios have approximately the same probability ratio of WO T2D $\approx 0.51$ . The effect of L on Death seems proportional to beta. 

```{r}
jpeg("T2D_10.jpg")
ggplot(melt(results10, measure.vars = 1:6))+
  geom_line(aes(x = i, y = value, col = variable), size = 1, linetype = "dashed")
dev.off()

ggplot(melt(dataa10, measure.vars = 1:6))+
  geom_line(aes(x = i, y = value, col = variable), size = 1, linetype = "dashed")
```
Same conclusion, just different numbers.

```{r}
results |> apply(2, FUN = mean)
results |> apply(2, FUN = sd)

results10 |> apply(2, FUN = mean)
results10 |> apply(2, FUN = sd)
```



